---
title: "1409 daily"
author: "Junsong Tang"
date: "5/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tseries)
library(xts)
library(zoo)
library(tidyverse)
library(forecast)
library(stlplus)
library(seasonal)

load("D:/NSERC/project/processedData/1409_dat.RData")

```

### data pre-processing


# encapsulated function form of daily data process
# input: response(covariate) matrix; boolean(True if impute N/A)
# output: data frame with year, month, day, & response(covariate) values
```{r, echo=TRUE}
dailyFrame = function(data, impute=FALSE) {
  # number of years
  n = length(data[, 1])
  
  # format unit month & days in calender
  month = NULL
  day = NULL
  for (m in (1:12)) {
    if ((m <= 7 && m %% 2 == 1) || (m >= 8 && m %% 2 == 0)) {
      month = c(month, rep(m, 31))
      day = c(day, seq(from = 1, to = 31, by = 1))
    } else if (m == 2) {
      month = c(month, rep(m, 28))
      day = c(day, seq(from = 1,to = 28,by = 1))
    } else {
      month = c(month, rep(m, 30))
      day = c(day, seq(from = 1,to = 30,by = 1))
    }
  }
  # repeat unit format for n years
  months = rep(month, n)
  days = rep(day, n)
  
  
  values = NULL
  years = NULL
  for (i in (1:n)) {
    for (j in (1:365)) {
      values = c(values, data[i,j])
      #N/A & imputation if applicable
      if (impute) {
        ind = (i-1)*365 + j
        if (is.na(values[ind])) {
          values[ind] = values[ind-365]
        }
      }
      years = c(years, 1980+i-1)
    }
  }
  
  return(data.frame(years, months, days, values)) 
}

```



#========================================================================================
# Daily unimputed data
```{r, echo=TRUE}
daily = dailyFrame(resp_mat)
# daily2 = data.frame(resp = as.vector(t(resp_mat)),
#                     rain = as.vector(t(rain_mat)),
#                     snow = as.vector(t(snow_mat)),
#                     temp = as.vector(t(temp_mat)),
#                     year = rep(1980:(1980+nrow(resp_mat)-1), each=ncol(resp_mat)),
#                     day = 1:ncol(resp_mat))

# identifying the range of NA
daily[is.na(daily$values),]

# subset of log-daily from 1980.1 - 1994.12(15 years)
daily_ts = ts(log(daily$values), start=c(1980,1), frequency = 365)
plot(daily_ts)
acf(daily_ts, lag.max = 365*5, main = "acf of daily response", na.action = na.pass)
pacf(daily_ts, lag.max = 365*5, main = "pacf of daily respoonse", na.action = na.pass)

# train set 12 years, test set 3 years
daily_train = window(daily_ts, start=c(1980,1), end=c(1991,365))
daily_test = window(daily_ts, start=c(1992,1), end=c(1994,365))


plot(daily_train, col="red", xlim = c(1980, 1995), main="Train & Test")
lines(daily_test, col="blue")
```




#========================================================================================
# Loess decomposition
```{r, echo=TRUE}
# loess = stlplus(daily_train, s.window = "periodic", s.degree = 2)
loess = stl(daily_train, s.window = "periodic")
plot(loess)
# check residuals
error = loess$time.series[,"remainder"]
qqnorm(error)
qqline(error)
acf(error, lag.max= 365*5)
pacf(error, lag.max = 365*3)
pacf(error)


# #de-seasonalize
# daily.train.ds = daily_train - loess$time.series[,"seasonal"]
# plot(daily.train.ds, main ="de-seasonalized")
# acf(daily.train.ds, lag.max = 365*5, main = "acf ds")
# pacf(daily.train.ds, lag.max = 365*5, main = "pacf ds")
# 
# # re-center
# daily.mu = mean(daily.train.ds)
# daily.train.rc = daily.train.ds - daily.mu
# plot(daily.train.rc, main ="de-seasonalized & re-centered ts")
# acf(daily.train.rc, lag.max = 365*5, main = "acf ds + rc")
# pacf(daily.train.rc, lag.max = 365*5, main = "pacf ds + rc")

```


# post-smoothing of seasonal
```{r, echo=TRUE}
rough_seasonal = seasonal(loess)
plot(rough_seasonal, type="l")


# simple_smooth = function(x, lag) {
#   n = length(x)
#   start = 1+lag
#   end = n-lag
#   result = NULL
#   for (i in (1:(end-start+1))) {
#     avg = mean(x[i:(i+2*lag)])
#     result = c(result, avg)
#   }
#   return (ts(result, frequency = 365))
# }

#window length = 1+lag
MA_smooth = function(x, lag) {
  n = length(x)
  result = NULL
  for (i in (1:(n-lag))) {
    avg = mean(x[i:(i+lag)])
    result = c(result, avg)
  }
  return (result)
}

ma_seas = MA_smooth(rough_seasonal, lag = 71)
plot(ma_seas, type="l")
lines(rough_seasonal, col="green")

spline_seas = smooth.spline(rough_seasonal, spar = 0.3)
plot(spline_seas, type = "l")
lines(rough_seasonal, col="green")



#shifting & padding of ma_seas
new_seas=NULL
for (i in (1:length(rough_seasonal))) {
  if (i <= 35 | i >= 4345) {
    new_seas[i] = NA
  } else {
    new_seas[i] = ma_seas[i-35]
  }
}

for (j in (1:length(rough_seasonal))) {
  if (j <= 35) {
    new_seas[j] = new_seas[j+365]
  } 
  if (j >= 4345) {
    new_seas[j] = new_seas[j-365]
  }
}

plot(new_seas, type="l")
lines(rough_seasonal, col="green")


```

# re-fit on residuals
```{r, echo=TRUE}
res = daily_train-loess$time.series[,"trend"]-ts(new_seas, start=c(1980,1), frequency = 365)
acf(res, lag.max = 365*10)
pacf(res, lag.max = 365*5)
pacf(res)

model = arima(res, order = c(5,0,0), include.mean = TRUE)
acf(model$residuals)
pacf(model$residuals, lag.max = 365)
tsdiag(model)

```



















<!-- # subseting function -->
<!-- ```{r, echo=TRUE} -->
<!-- # subseting clusters of ts values at given set of certain days within a period -->
<!-- same_days = function(ts, dayNum) { -->
<!--   # number of years -->
<!--   k = floor(tsp(ts)[2]-tsp(ts)[1])+1 -->
<!--   # inherent frequency of given time series -->
<!--   freq = as.integer(tsp(ts)[3]) -->

<!--   # sanity check -->
<!--   for (d in dayNum) { -->
<!--     if (d > freq || d <= 0) { -->
<!--       print("Day number input error!") -->
<!--       return () -->
<!--     } -->
<!--   } -->




<!-- } -->


<!-- ``` -->



#========================================================================================
Seasonality

1st version ($c = E[(A^TA)^{-1}A^TY] = (A^TA)^{-1}A^T E(Y)$)
```{r, echo=TRUE}
# (1) fit model to every member of same day class, then take expectation of fitted values at interval center 
season1 = function(ts, window, degree) {
  # number of years
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1
  freq = as.integer(tsp(ts)[3])
  
  seasonality = NULL
  
  for (i in (1:freq)) {
    same_days = NULL
    centers = NULL
    for (j in (1:k)) {
      center = i+(j-1)*freq
      
      lb = max(center-window, 1)
      ub = min(center+window, k*freq)
      idx = seq(from=lb,to=ub,by=1)
      x = idx
      y = ts[idx]
      
      # local regression
      if (degree == 0) {
        model = lm(y ~ 1)
        yhat = as.numeric(model$fitted.values[which(x==center)])
      }
      if (degree == 1) {
        model = lm(y ~ x)
        yhat = as.numeric(model$fitted.values[which(x==center)])
      }
      if (degree == 2) {
        model = lm(y ~ x+I(x^2))
        yhat = as.numeric(model$fitted.values[which(x==center)])
      }
      if (degree == 3) {
        model = lm(y ~ x+I(x^2)+I(x^3))
        yhat = as.numeric(model$fitted.values[which(x==center)])
      }
      
      
      #collect same day values
      same_days = c(same_days, yhat)
      
      centers = c(centers, center)
    }
    
    # unify all same day values
    day_level = mean(same_days)
    
    # slot = same day index in the original series
    # slots = seq(from=i,to=i+(k-1)*freq, by=freq)
    seasonality[centers] = day_level
  }
  
  return (seasonality)
  
}

#========================================================================================
# (2) collect all y at same day, take expectation of each cluster(y), fit a model, and take fitted values at interval center.
season2 = function(ts, window, degree) {
  # number of years
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1
  freq = as.integer(tsp(ts)[3])
  
  seasonality = NULL
  
  for (i in (1:freq)) {
    # collection of y at 
    same_days = NULL
    
    interval = seq(from=-window+i, to=window+i, by=1)
    
    for (w in interval) {
      indices = NULL
      for (j in (0:k)) {
        idx = w + j*freq
        if (1 <= idx && idx <= k*freq) {
          indices = c(indices, idx)
        }
      }
      
      # same day cluster
      cluster = ts[indices]
      m = mean(cluster)
      
      same_days = c(same_days, m)
    }
    
    y = same_days
    x = interval
    
    if (degree == 0) {
      model = lm(y ~ 1)
      yhat = as.numeric(model$fitted.values[which(x==i)])
    }
    if (degree == 1) {
      model = lm(y ~ x)
      yhat = as.numeric(model$fitted.values[which(x==i)])
    }
    if (degree == 2) {
      model = lm(y ~ x+I(x^2))
      yhat = as.numeric(model$fitted.values[which(x==i)])
    }
    if (degree == 3) {
      model = lm(y ~ x+I(x^2)+I(x^3))
      yhat = as.numeric(model$fitted.values[which(x==i)])
    }
    
    slots = seq(from=i, to=i+(k-1)*freq, by=freq)
    seasonality[slots] = yhat
  }
  
  return (seasonality)
}


```









2nd version
# (1) x = (-2,-1,0,1,2,-2,-1,0,1,2,...); y = ($,$,$,$,$,&,&,&,&,&,...) horizontal then vertical(this method is not convenient to remove outliers in each cluster at same day)
```{r, echo=TRUE}
seasonality1 = function(ts, window, degree) {
  # number of years
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1
  freq = as.integer(tsp(ts)[3])
  
  seasonality = NULL
  
  for (i in (1:freq)) {
    centers = NULL
    y = NULL
    x = NULL
    for (j in (1:k)) {
      center = i+(j-1)*freq
      
      lb = max(center-window, 1)
      ub = min(center+window, k*freq)
      indices = seq(from=lb,to=ub,by=1)
      normalize_indices = indices - center
      
      
      level = ts[indices]
      # remove x whose corresponding y is NA
      normalize_indices = normalize_indices[!is.na(level)]
      # remove NA in y
      level = level[!is.na(level)]
      
      y = c(y, level)
      x = c(x, normalize_indices)
      
      centers = c(centers, center)
    }
    
    # local regression
    if (degree == 0) {
      model = lm(y ~ 1)
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 1) {
      model = lm(y ~ x)
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 2) {
      model = lm(y ~ x+I(x^2))
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 3) {
      model = lm(y ~ x+I(x^2)+I(x^3))
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    
    # slot = same day index in the original series
    # slots = seq(from=i,to=i+(k-1)*freq, by=freq)
    seasonality[centers] = yhat
  }
  return (seasonality)
}

```
# (2) x = (-2,-2,-2,...,-1,-1,-1,...,0,0,0,...,1,1,1,...,2,2,2,...); 
# y = ($,&,#,...,$,&,#,...,$,&,#,...,$,&,#,...,$,&,#,...,), vertical then horizontal
```{r, echo=TRUE}
seasonality2 = function(ts, window, degree) {
  # number of years
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1
  freq = as.integer(tsp(ts)[3])
  
  seasonality = NULL
  
  for (i in (1:freq)) {
    # [-w+i,..i,...w+i]
    interval = seq(from=-window+i, to=window+i, by=1)
    
    y = NULL
    x = NULL
    
    for (w in interval) {
      indices = NULL
      for (j in (1:k)) {
        idx = w + (j-1)*freq
        # check if valid index
        if (1 <= idx && idx <= k*freq) {
          indices = c(indices, idx)
          # if (!is.na(ts[idx])) {
          #   x = c(x, w-i)
          #   cluster = c(cluster, ts[idx])
          # }
        }
      }
      
      # same day cluster
      cluster = ts[indices]
      # remove NA in cluster
      cluster = cluster[!is.na(cluster)]
      
      y = c(y, cluster)
      # normalized interval center is always 0:[-w,...0,...w]
      x = c(x, rep(w-i, length(cluster)))
      
    }
    
    if (degree == 0) {
      model = lm(y ~ 1)
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 1) {
      model = lm(y ~ x)
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 2) {
      model = lm(y ~ x+I(x^2))
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 3) {
      model = lm(y ~ x+I(x^2)+I(x^3))
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    
    slots = seq(from=i, to=i+(k-1)*freq, by=freq)
    seasonality[slots] = yhat
  }
  
  return (seasonality)
}

```

#(3) remove outliers and fit regression model
```{r, echo=TRUE}
seasonality3 = function(ts, window, degree) {
  # number of years
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1
  freq = as.integer(tsp(ts)[3])
  
  seasonality = NULL
  
  for (i in (1:freq)) {
    # [-w+i,..i,...w+i]
    interval = seq(from=-window+i, to=window+i, by=1)
    
    y = NULL
    x = NULL
    
    for (w in interval) {
      indices = NULL
      # cluster = NULL
      for (j in (1:k)) {
        idx = w + (j-1)*freq
        # check if valid index
        if (1 <= idx && idx <= k*freq) {
          indices = c(indices, idx)
          # if (!is.na(ts[idx])) {
          #   x = c(x, w-i)
          #   cluster = c(cluster, ts[idx])
          # }
        }
      }
      
      # same day cluster
      cluster = ts[indices]
      # remove NA in cluster
      cluster = cluster[!is.na(cluster)]
      # set quantile & subset cluster without outlier  and subset x accordingly s.t same_x, cluster have same length
      # range is a vector: c(lb, ub)
      range = quantile(cluster, prob = c(0.05, 0.95))
      lb = as.numeric(range[1])
      ub = as.numeric(range[2])
      cluster = cluster[lb<=cluster&cluster<=ub]

      y = c(y, cluster)
      # normalized interval center is always 0:[-w,...0,...w]
      x = c(x, rep(w-i, length(cluster)))
    }
    
    if (degree == 0) {
      model = lm(y ~ 1)
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 1) {
      model = lm(y ~ x)
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 2) {
      model = lm(y ~ x+I(x^2))
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 3) {
      model = lm(y ~ x+I(x^2)+I(x^3))
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    
    slots = seq(from=i, to=i+(k-1)*freq, by=freq)
    seasonality[slots] = yhat
  }
  
  return (seasonality)
}


```



# test using 2nd version
```{r, echo=TRUE}
start_time = Sys.time()
s1 = seasonality1(daily_ts, window = 51, degree = 2)
s2 = seasonality2(daily_ts, window = 51, degree = 2)
s3 = seasonality3(daily_ts, window = 51, degree = 2)
end_time = Sys.time()
end_time - start_time

```


# plot seasonality
```{r, echo=TRUE}
plot(s1, type ="l", xlim = c(0,1000), col = "black")
lines(s2, type ="l", xlim = c(0,1000), col = "red")
# lines(s3, type ="l", xlim = c(0,1000), col = "green")
```


# runtime for method 1
```{r, echo=TRUE}
start_time = Sys.time()
s1 = seasonality1(daily_ts, window = 51, degree = 2)
end_time = Sys.time()
end_time - start_time

```

# runtime for method 2
```{r, echo=TRUE}
start_time = Sys.time()
s2 = seasonality2(daily_ts, window = 51, degree = 2)
plot(s2, type ="l", xlim = c(0,1000))
end_time = Sys.time()
end_time - start_time

```

# runtime for method 3
```{r, echo=TRUE}
start_time = Sys.time()
s3 = seasonality3(daily_ts, window = 51, degree = 2)
plot(s3, type ="l", xlim = c(0,1000))
end_time = Sys.time()
end_time - start_time

```


```{r, echo=TRUE}
remain_loess = stl(daily_ts-s3, s.window = "periodic")
plot(remain_loess)


```





















# test using 1st version
```{r, echo=TRUE}
start_time = Sys.time()
seas1 = season1(daily_train, window = 51, degree = 2)
plot(seas1, type ="l")
end_time = Sys.time()
end_time - start_time


# daily_ds = daily_train-seas
# decomp = stl(daily_ds, s.window = "periodic")
# plot(decomp)

```



```{r, echo=TRUE}
start_time = Sys.time()
seas2 = season2(daily_train, window = 51, degree = 2)
plot(seas2, type ="l")
end_time = Sys.time()
end_time - start_time
```












#========================================================================================
# Model Fitting
<!-- # Classical Decomposition by moving average -->
<!-- ```{r, echo=TRUE} -->
<!-- component = decompose(daily_train, type = "additive", filter = c()) -->
<!-- plot(component) -->
<!-- acf(component$random, na.action = na.pass, lag.max = 365*10) -->
<!-- pacf(component$random, na.action = na.pass, lag.max = 365*5) -->
<!-- pacf(component$random, na.action = na.pass) -->

<!-- ``` -->

# ARIMA(5,0,0) model fitting
```{r, echo=TRUE}
arima_model = arima(error, order = c(5,0,0), include.mean = TRUE)
plot(arima_model$residuals)
acf(arima_model$residuals)
pacf(arima_model$residuals, lag.max = 365)
tsdiag(arima_model)
qqnorm(arima_model$residuals)
qqline(arima_model$residuals)

```


# Box-Jenkins Forecast Method Function
```{r, echo=TRUE}
box_jenkins = function(model, ahead) {
  n = length(model$coef)
  mu = model$coef[n]
  preds = NULL
  order = n - 1
  # the list of previous values till today
  queue = window(error, start=c(1991,365-order+1), end=c(1991,365))
  while (ahead != 0) {
    pred = 0
    for (i in (1:order)) {
      pred = pred + as.numeric(queue[order-i+1]) * as.numeric(model$coef[i])
    }
    preds = c(preds, pred)
    queue = c(queue[-1], pred)
    ahead = ahead - 1
  }
  return (ts(preds, start=c(1992,1), frequency = 365))
  
}


```



# ARIMA(5,0,0) model forecast
```{r, echo=TRUE}
lag = 15
future_days = window(daily_test, start=c(1992,1), end=c(1992,lag))


pred1 = predict(arima_model, n.ahead = lag)
predict_values = pred1$pred + loess$time.series[, "trend"][1:lag] + loess$time.series[,"seasonal"][1:lag]
plot(predict_values, col="black", ylim = c(-1.5,-0.5), main="prediction")
lines(future_days, col="red")

#box-jenkins method prediction
pred2 = box_jenkins(arima_model, ahead = lag)
predict_values2 = pred2 + loess$time.series[, "trend"][1:lag] + loess$time.series[,"seasonal"][1:lag]
plot(predict_values2, col="black", ylim = c(-1.5,-0.5), main="prediction")
lines(future_days, col="red")


plot(pred1$pred, col="black", main="prediction")
lines(pred2, col="red")


```




























#========================================================================================
# Fourier method to decompose long seasonality(p=365)
```{r, echo=TRUE}

fourier_comp = function(Xt, order) {
  fc = fourier(Xt, K=order)
  St = NULL
  for (t in (1:length(Xt))) {
    s = 0 # fourier partial sum at time t
    for (k in (1:order)) {
      sin_term = fc[t, 2*k-1]
      cos_term = fc[t, 2*k]
      # s = s + a * sin(((2*pi*k)/365) * t) + b * cos(((2*pi*k)/365) * t)
      s = s + sin_term + cos_term
    }
    St = c(St, s)
  }
  return (St)
}

St = fourier_comp(daily_train, order = 2)
plot(daily_train, main="Daily train set")
plot(St, main="seasonal component")
plot(daily_train-St, main = "residuals")


curr_aic = Inf
order = 1
for (k in (1:4)) {
  model = auto.arima(daily_train, xreg = fourier(daily_train, K=k))
  if (model$aic < curr_aic) {
    curr_aic = model$aic
    order = k
  }
}




daily_model = auto.arima(daily_train, xreg = fourier(daily_train, K=2))

tsdiag(daily_model)
qqnorm(daily_model$residuals)
qqline(daily_model$residuals)


```





# forecast
```{r, echo=TRUE}
# 1st prediction, forecast function
plot(forecast(daily_model, h=3*365, xreg=fourier(daily_train, K=2, h=3*365)))
lines(daily_test, col="red")
legend("bottomleft",legend=c("predict value","actual value"), lty=1, col=c("blue","red"))

# 2nd prediction
pred = predict(daily_model, newxreg=fourier(daily_train, K=2, h=3*365), n.ahead = 3*365)
plot(pred$pred, col="black", ylim = c(-1,2), main="prediction")
lines(daily_test, col="red")


```


# model 2 & forecast 2
``` {r, echo=TRUE}
model2 = auto.arima(daily_train, xreg = fourier(daily_train, K=3))

tsdiag(model2)
qqnorm(model2$residuals)
qqline(model2$residuals)

plot(forecast(model2, h=3*365, xreg=fourier(daily_train, K=3, h=3*365)))
lines(daily_test, col="red")
legend("bottomleft",legend=c("predict value","actual value"), lty=1, col=c("blue","red"))

# 2nd prediction

plot(predict(model2, newxreg=fourier(daily_train, K=3, h=3*365), n.ahead = 3*365)$pred, col="black", ylim = c(-1,2), main="prediction")
lines(daily_test, col="red")


```

```{r, echo=TRUE}
RMSE2 = sqrt(mean((pred$pred - daily_test)^2))
RMSE3 = sqrt(mean((predict(model2, newxreg=fourier(daily_train, K=3, h=3*365), n.ahead = 3*365)$pred - daily_test)^2))
```



# constraint on MA order
```{r,echo=TRUE}
model = auto.arima(daily_train, max.q = 0, max.Q = 0, xreg = fourier(daily_train, K=3))


tsdiag(model)
qqnorm(model$residuals)
qqline(model$residuals)

plot(forecast(model, h=3*365, xreg=fourier(daily_train, K=3, h=3*365)))
lines(daily_test, col="red")
legend("bottomleft",legend=c("predict value","actual value"), lty=1, col=c("blue","red"))


plot(predict(model, newxreg=fourier(daily_train, K=3, h=3*365), n.ahead = 3*365)$pred, col="black", ylim = c(-1,2), main="prediction")
lines(daily_test, col="red")

# sqrt(mean((predict(model, newxreg=fourier(daily_train, K=3, h=3*365), n.ahead = 3*365)$pred - daily_test)^2))
```




```{r, echo=TRUE}
fourier_model = function(Xt) {
  models = NULL
  curr_aic = Inf
  order = 1
  for (k in (1:4)) {
    model = auto.arima(Xt, max.q = 0, max.Q = 0, xreg = fourier(Xt, K=k))
    models = c(models, model)
    if (model$aic < curr_aic) {
      curr_aic = model$aic
      order = k
    }
  }
  return (models[order])
}

```


```{r, echo=TRUE}
modell = auto.arima(daily_train, max.q = 0, max.Q = 0, max.d = 0, max.D = 0, xreg = fourier(daily_train, K=2))


tsdiag(modell)
qqnorm(modell$residuals)
qqline(modell$residuals)

plot(forecast(modell, h=3*365, xreg=fourier(daily_train, K=2, h=3*365)))
lines(daily_test, col="red")
legend("bottomleft",legend=c("predict value","actual value"), lty=1, col=c("blue","red"))


plot(predict(modell, newxreg=fourier(daily_train, K=2, h=3*365), n.ahead = 3*365)$pred, col="black", ylim = c(-1,2), main="prediction")
lines(daily_test, col="red")

```


