---
title: "1409 daily"
author: "Junsong Tang"
date: "5/16/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tseries)
library(xts)
library(zoo)
library(tidyverse)
library(forecast)
library(stlplus)
library(seasonal)
library(L1pack)


```


<!-- ```{r, echo=TRUE} -->
<!-- dailyFrame = function(data, impute=FALSE) { -->
<!--   # number of years -->
<!--   n = length(data[, 1]) -->

<!--   # format unit month & days in calender -->
<!--   month = NULL -->
<!--   day = NULL -->
<!--   for (m in (1:12)) { -->
<!--     if ((m <= 7 && m %% 2 == 1) || (m >= 8 && m %% 2 == 0)) { -->
<!--       month = c(month, rep(m, 31)) -->
<!--       day = c(day, seq(from = 1, to = 31, by = 1)) -->
<!--     } else if (m == 2) { -->
<!--       month = c(month, rep(m, 28)) -->
<!--       day = c(day, seq(from = 1,to = 28,by = 1)) -->
<!--     } else { -->
<!--       month = c(month, rep(m, 30)) -->
<!--       day = c(day, seq(from = 1,to = 30,by = 1)) -->
<!--     } -->
<!--   } -->
<!--   # repeat unit format for n years -->
<!--   months = rep(month, n) -->
<!--   days = rep(day, n) -->


<!--   values = NULL -->
<!--   years = NULL -->
<!--   for (i in (1:n)) { -->
<!--     for (j in (1:365)) { -->
<!--       values = c(values, data[i,j]) -->
<!--       #N/A & imputation if applicable -->
<!--       if (impute) { -->
<!--         ind = (i-1)*365 + j -->
<!--         if (is.na(values[ind])) { -->
<!--           values[ind] = values[ind-365] -->
<!--         } -->
<!--       } -->
<!--       years = c(years, 1980+i-1) -->
<!--     } -->
<!--   } -->

<!--   return(data.frame(years, months, days, values)) -->
<!-- } -->

<!-- ``` -->


#=======================================================================================
# Daily unimputed data
```{r, echo=TRUE}
# data1409 = mget(load("D:/NSERC/project/processedData/1409_dat.RData"))
daily = daily_frame(data1409)
# daily = dailyFrame(data1409$resp_mat)
# daily2 = data.frame(resp = as.vector(t(resp_mat)),
#                     rain = as.vector(t(rain_mat)),
#                     snow = as.vector(t(snow_mat)),
#                     temp = as.vector(t(temp_mat)),
#                     year = rep(1980:(1980+nrow(resp_mat)-1), each=ncol(resp_mat)),
#                     day = 1:ncol(resp_mat))

# identifying the range of NA
daily[is.na(daily$resp),]

# subset of log-daily from 1980.1 - 1994.12(15 years)
ts1409 = ts(log(daily$resp), start=c(1980,1), frequency = 365)
plot(ts1409)
acf(ts1409, lag.max = 365*5, main = "acf of daily response", na.action = na.pass)
pacf(ts1409, lag.max = 365*5, main = "pacf of daily respoonse", na.action = na.pass)

# # train set 12 years, test set 3 years
# daily_train = window(daily_ts, start=c(1980,1), end=c(1991,365))
# daily_test = window(daily_ts, start=c(1992,1), end=c(1994,365))


# plot(daily_train, col="red", xlim = c(1980, 1995), main="Train & Test")
# lines(daily_test, col="blue")
```




#========================================================================================
# STL decomposition
```{r echo=TRUE}
# stl_obj = stlplus(daily_train, s.window = "periodic", s.degree = 2)
na.new = function(x) {ts(zoo::na.approx(x), start = c(1980,1), frequency = 365)}
stl_obj = stl(ts1409, s.window = "periodic", na.action = na.new)
plot(stl_obj)
# check residuals
# daily_ts has length 13870; while residual has 13398, probably due to NAs in raw ts.
error = stl_obj$time.series[,"remainder"]
qqnorm(error)
qqline(error)
acf(error, lag.max= 365*5, main="acf of residual")
pacf(error, lag.max = 365*3, main="pacf of residual")
pacf(error, main="pacf of residual")


model = arima(error, order = c(5,0,0), include.mean = TRUE)
acf(model$residuals)
pacf(model$residuals, lag.max = 365)
tsdiag(model)
qqnorm(model$residuals)
qqline(model$residuals)
hist(model$residuals)

# #de-seasonalize
# daily.train.ds = daily_train - loess$time.series[,"seasonal"]
# plot(daily.train.ds, main ="de-seasonalized")
# acf(daily.train.ds, lag.max = 365*5, main = "acf ds")
# pacf(daily.train.ds, lag.max = 365*5, main = "pacf ds")
# 
# # re-center
# daily.mu = mean(daily.train.ds)
# daily.train.rc = daily.train.ds - daily.mu
# plot(daily.train.rc, main ="de-seasonalized & re-centered ts")
# acf(daily.train.rc, lag.max = 365*5, main = "acf ds + rc")
# pacf(daily.train.rc, lag.max = 365*5, main = "pacf ds + rc")

```










# My method
#=========================================================================================
# Seasonality

1st version ($c = E[(A^TA)^{-1}A^TY] = (A^TA)^{-1}A^T E(Y)$)

2nd version
# (1) x = (-2,-1,0,1,2,-2,-1,0,1,2,...); y = ($,$,$,$,$,&,&,&,&,&,...) horizontal then vertical(this method is not convenient to remove outliers in each cluster at same day)

# (2) x = (-2,-2,-2,...,-1,-1,-1,...,0,0,0,...,1,1,1,...,2,2,2,...); 
# y = ($,&,#,...,$,&,#,...,$,&,#,...,$,&,#,...,$,&,#,...,), vertical then horizontal


# helper: re-arrange the raw time series into same day equivalent classes
```{r, echo=TRUE}
# trim: boolean - whether to trim out the outliers of cluster(matrix row)
# same_day_cluster = function(ts, trim=FALSE) {
#   k = floor(tsp(ts)[2]-tsp(ts)[1])+1
#   freq = as.integer(tsp(ts)[3])
# 
#   M = list()
#   for (i in (1:freq)) {
#     indices = seq(from=i, to=i+(k-1)*freq, by=freq)
#     row = ts[indices]
#     # whether to remove outliers or not
#     if (trim) {
#       # set quantile & subset
#       range = quantile(row, prob = c(0.25, 0.75), na.rm = TRUE)
#       lb = as.numeric(range[1])
#       ub = as.numeric(range[2])
#       row = row[lb<=row&row<=ub]
#     }
#     M[[i]] = row
#   }
#   return (M)
# }


# trim: percent of lower tail
same_day_cluster = function(ts, trim=0) {
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1
  freq = as.integer(tsp(ts)[3])

  M = list()
  for (i in (1:freq)) {
    indices = seq(from=i, to=i+(k-1)*freq, by=freq)
    row = ts[indices]
    # whether to remove outliers or not
    if (trim != 0) {
      # set quantile & subset
      range = quantile(row, prob = c(trim, 1-trim), na.rm = TRUE)
      lb = as.numeric(range[1])
      ub = as.numeric(range[2])
      row = row[lb<=row&row<=ub]
    }
    M[[i]] = row
  }
  return (M)
}




```




```{r, echo=TRUE}
# trim: real number in [0,1], representing percentage of the lower tail
seasonality = function(ts, window, degree, trim) {
  # number of years
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1
  freq = as.integer(tsp(ts)[3])
  start_year = as.integer(tsp(ts)[1])
  
  equi_class = same_day_cluster(ts, trim=trim)
  n = length(equi_class[[1]])
  
  seasonality = NULL
  for (i in (1:freq)) {
    # [-w+i,..i,...w+i]
    interval = seq(from=-window+i, to=window+i, by=1)
    y = NULL
    x = NULL
    for (w in interval) {
      if (w < 1) {
        cluster = (equi_class[[w+freq]])[-n]
      }
      else if (w > freq) {
        cluster = (equi_class[[w-freq]])[-1]
      }
      else {
        cluster = equi_class[[w]]
      }
      
      cluster = cluster[!is.na(cluster)]
      
      y = c(y, cluster)
      x = c(x, rep(w-i, length(cluster)))
    }
    
    if (degree == 0) {
      model = lm(y ~ 1)
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 1) {
      model = lm(y ~ x)
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 2) {
      model = lm(y ~ x+I(x^2))
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 3) {
      model = lm(y ~ x+I(x^2)+I(x^3))
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    
    slots = seq(from=i, to=i+(k-1)*freq, by=freq)
    seasonality[slots] = yhat
  }
  
  return (ts(seasonality, start=c(start_year, 1), frequency = freq))
}
```


# run time test using 2nd version

# runtime for seasonality
```{r, echo=TRUE}
start_time = Sys.time()
# keep the outliers
s1 = seasonality(ts1409, window = 51, degree = 2, trim=0)
end_time = Sys.time()
end_time - start_time

```

```{r, echo=TRUE}
start_time = Sys.time()
# remove the outliers
s2 = seasonality(ts1409, window = 51, degree = 2, trim=0.25)
end_time = Sys.time()
end_time - start_time

```






# plot seasonality
```{r, echo=TRUE}
plot(s1, type ="l", xlim = c(1980,1983), col = "black", main="s1&s2: with/without outliers")
lines(s2, type = "l", xlim = c(1980, 1983), col = "red")
legend("bottomright",legend=c("outlier","no outlier"), lty=1,
       col=c("black","red"))


```








# Cross validation(5 fold by default)
# 38 => 8 + 8 + 8 + 8 + 6
```{r, echo=TRUE}
cv.myseason = function(ts, window.max = 10, trim_per.max = 0.25, fold = 5) {
  start_year = tsp(ts)[1]
  end_year = tsp(ts)[2]
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1 #k = 38 years
  t = ceiling(k/fold) #t = 8 years/unit
  freq = as.integer(tsp(ts)[3])
  minSMSR = Inf
  best_window = 0
  best_degree = 0
  best_trim = 0
  for (w in (0:window.max)) {
    for (d in (0:3)) {
      for (p in (0:(trim_per.max/0.05))) {
        # sum of mean squared error over CVs
        SMSR = 0
        # run fold times CV
        for (i in (1:fold)) {
          test_index = seq(from=1+(i-1)*t*freq, to=min(i*t*freq, k*freq), by=1)
          test = ts[test_index]
          test = ts(test, start=c(start_year+(i-1)*t, 1), frequency = freq)
          
          train = ts[-test_index]
          train = ts(train, start=c(start_year, 1), frequency = freq)
          seasonality = seasonality(train, w, d, trim=p*0.05)
          
          
          pred = seasonality[1:length(test)]
          residual = (test - pred)
          SMSR = SMSR + mean(residual^2, na.rm = TRUE)
        }
        
        
        
        # for (year in (1:k)) {
        #   test_index = seq(from=1+(year-1)*freq, to=year*freq, by=1)
        #   test = ts[test_index]
        #   # test must be converted into ts object!!!
        #   test = ts(test, start=c(start_year+(year-1), 1), frequency = freq)
        #   train = ts[-test_index]
        #   # train must be converted into ts object!!!
        #   train = ts(train, start=c(start_year, 1), frequency = freq)
        #   seasonality = seasonality4(train, w, d)
        #   pred = seasonality[1:freq]
        #   residual = (test - pred)
        #   SMSR = SMSR + mean(residual^2, na.rm = TRUE)
        # }
        
        if (SMSR < minSMSR) {
          best_window = w
          best_degree = d
          best_trim = p
          minSMSR = SMSR
        }
      }
    }
  }
  
  return (data.frame(best_window, best_degree, best_trim))
}

```




# test
```{r, echo=TRUE}
start_time = Sys.time()
s_opt = seasonality(ts1409, window=11, degree=0, trim=0.05)
end_time = Sys.time()
end_time - start_time
plot(ts1409, col = "green", xlim=c(1980,2000), ylim=c(-5,5), type="l", main="best seasonality(11, 0, 5%) & raw ts")
lines(s_opt, col = "blue")
lines(loess$time.series[,"seasonal"], col = "red")
legend("bottomright",legend=c("ts","mine", "stl"), lty=1,
       col=c("green","blue","red"))

# remain = daily_ts - best_seasonal
# obj = stl(remain, s.window = "periodic", na.action = na.pass)
# plot(obj)

```








#==========================================================================================
# Trend

(1) moving average over period(center at t=183)
```{r, echo=TRUE}
MA.smooth = function(x, lag) {
  n = length(x)
  result = NULL
  for (i in (1:(n-lag))) {
    avg = mean(x[i:(i+lag)], na.rm = TRUE)
    result = c(result, avg)
  }
  return (result)
}


trend = function(ts) {
  freq = as.integer(tsp(ts)[3])
  n = length(ts)
  start_year = as.integer(tsp(ts)[1])
  
  trend = NULL
  final_trend = rep(NA, n)
  # freq=5, w=2, freq=4, w=1.5
  window  = (freq-1) / 2
  # if freq = odd
  if (freq %% 2 != 0) {
    trend = MA.smooth(ts, lag=2*window)
    # padding head & tail
    final_trend[(window+1):(n-window)] = trend
  }
  else {
    temp = MA.smooth(ts, lag=2*window)
    trend = MA.smooth(temp, lag = 1)
    # padding head & tail
    w = window + 0.5
    final_trend[(w+1):(n-w)] = trend
  }
  
  return (ts(final_trend, start=c(start_year, 1), frequency = freq))
  
}


```


# test
```{r, echo=TRUE}
start_time = Sys.time()
t1 = trend(ts1409)
end_time = Sys.time()
end_time - start_time
plot(ts1409, ylim = c(-5,5), type = "l", col = "green", main="trend in ts")
lines(t1, col = "blue")
lines(stl_obj$time.series[,"trend"], col = "red")
legend("bottomright",legend=c("ts","mine", "stl"), lty=1,
       col=c("green","blue","red"))
```




```{r, echo=TRUE}
res = ts1409 - s_opt - t1

qqnorm(res)
qqline(res)
par(mar=c(5,4,4,2)+0.1)
acf(res[!is.na(res)], lag.max= 365*5, main="acf of residual")
pacf(res[!is.na(res)], lag.max = 365*3, main="pacf of residual")
pacf(res[!is.na(res)], main="pacf of residual")


# model = arima(res, order = c(5,0,0), include.mean = TRUE)
# acf(model$residuals)
# pacf(model$residuals, lag.max = 365)
tsdiag(arima(res[!is.na(res)], order = c(5,0,0), include.mean = TRUE))

hist(error, main="residual by stl")


```















# LOESS
#==========================================================================================
# loess method

#preparation
```{r, echo=TRUE}
# NA removal version
same_day_frame = function(ts) {
  # number of years
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1
  freq = as.integer(tsp(ts)[3])

  x = NULL
  y = NULL
  for (i in (1:freq)) {
    indices = seq(from=i, to=i+(k-1)*freq, by=freq)
    cluster = ts[indices]
    # clear out NA
    cluster = cluster[!is.na(cluster)]

    x = c(x, rep(i, length(cluster)))
    y = c(y, cluster)
    
  }
  df = data.frame(x, y)
  return (df)
}

## NA kept version
# same_day_frame = function(ts) {
#   # number of years
#   k = floor(tsp(ts)[2]-tsp(ts)[1])+1
#   freq = as.integer(tsp(ts)[3])
#   
#   m = matrix(ts, byrow = TRUE, nrow=k)
#   x = rep(1:freq, each = length(m[,1]))
#   y = as.vector(m)
#   return (data.frame(x, y))
# }


```

```{r, echo=TRUE}
start_time = Sys.time()
df = same_day_frame(ts1409)
end_time = Sys.time()
end_time - start_time


```


<!-- ```{r, echo=TRUE} -->
<!-- start_time = Sys.time() -->
<!-- s = loess(y ~ x, data=df, degree=2, family="symmetric", method="loess", control =loess.control(iterations = 5)) -->
<!-- end_time = Sys.time() -->
<!-- end_time - start_time -->


<!-- ``` -->

#seasonality
```{r, echo=TRUE}
# s = loess(y ~ x, data=df, degree=2, family="symmetric", method="loess", control =loess.control(iterations = 4))
# # j <- order(df$x)
# # plot(df$y ~ df$x)
# # lines(df$x[j], s$fitted[j], col="green")
# fitted_frame = data.frame(df$x, s$fitted)
# names(fitted_frame) = c("x", "fitted")
# 
# count = 1
# fitted = NULL
# for (i in (1:length(fitted_frame$x))) {
#   if (fitted_frame$x[i] == count && count <= 365) {
#     fitted = c(fitted, fitted_frame$fitted[i])
#     count = count + 1
#   }
# }
# plot(df$y ~ df$x)
# lines(seq(from=1, to=365, by=1), fitted, col="green")
# plot(best_seasonal[1:365], type = "l", xlim = c(0, 365), ylim = c(-5, 2))
# lines(fitted, col = "red")

```


```{r, echo=TRUE}
seasonal.loess = function(ts, deg=2, span=0.75, iter=4) {
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1
  freq = as.integer(tsp(ts)[3])
  start_year = as.integer(tsp(ts)[1])
  # form data frame arranged in same day order
  x = NULL
  y = NULL
  z = NULL
  for (i in (1:freq)) {
    indices = seq(from=i, to=i+(k-1)*freq, by=freq)
    cluster = ts[indices]
    # clear out NA
    cluster = cluster[!is.na(cluster)]
    x = c(x, rep(i, length(cluster)))
    y = c(y, cluster)
    # set up marker such that it represents the pos that we should extract the fitted value 
    z = c(z, TRUE, rep(FALSE, length(cluster)-1))

  }
  # df = data.frame(x, y)

  loess_obj = loess(y ~ x, span=span, degree=deg, family="symmetric", method="loess", control =loess.control(iterations=iter))
  
  full_period = loess_obj$fitted[which(z==TRUE)]
  seasonality = rep(full_period, k)
  
  return (ts(seasonality, start=c(start_year,1), frequency = freq))
}



```



#test
```{r, echo=TRUE}
start_time = Sys.time()
s3 = seasonal.loess(ts1409, span=1/30)
end_time = Sys.time()
end_time - start_time
plot(df$y ~ df$x, col = "green", main = "one full period")
lines(s_opt[1:365], col = "red")
lines(s3[1:365], col = "blue")
legend("topright",legend=c("ts","loess", "mine"), lty=1,
       col=c("green","blue","red"))
```

```{r, echo=TRUE}
plot(ts1409, type="l", xlim = c(1980,1982), ylim=c(-3,3), col="green")
lines(s3, col="blue")
lines(s_opt, col="red")
```








#trend

```{r, echo=TRUE}
trend.loess = function(ts, deg=2, span=0.75, iter=4) {
  freq = as.integer(tsp(ts)[3])
  start_year = as.integer(tsp(ts)[1])
  n = length(ts)
  window_ref = n * span
  trend = NULL
  # 1st cursor, controlling pos in whole ts
  start_pos = 1
  while (start_pos <= n) {
    x = NULL
    y = NULL
    # 2nd cursor, controlling accumulating consecutive non-NA
    i = start_pos
    while (!is.na(ts[i])) {
      y = c(y, ts[i])
      x = c(x, i)
      i = i + 1
    }

    # first time meets NA after consecutive strings of non-NA
    if (!is.null(y)) {
      new_span = window_ref / length(y)
      loess_obj = loess(y ~ x, span=new_span, degree=deg, family="symmetric", method="loess", control=loess.control(surface="direct", iterations=iter))
      trend = c(trend, loess_obj$fitted)
      start_pos = i
    }
    
    # if (!is.null(y) && length(y) >= window) {
    #   loess_obj = loess(y ~ x, span=span, degree=deg, family="symmetric", method="loess", control=loess.control(surface="direct", iterations=iter))
    #   trend = c(trend, loess_obj$fitted)
    #   if (i <= n) {
    #     trend = c(trend, NA)
    #   }
    # }
    # # some non-NA but length too short
    # else if (length(y) < window) {
    #   trend = c(trend, rep(NA, length(y)))
    #   if (i <= n) {
    #     trend = c(trend, NA)
    #   }
    # }
    
    # consecutive NAs
    else {
      trend = c(trend, NA)
      start_pos = i + 1
    }
  }

  return (ts(trend, start=c(start_year, 1), frequency = freq))
}



```


<!-- ```{r, echo=TRUE} -->
<!-- trend.loess = function(ts, deg=2, span=0.75, iter=4, impute=FALSE) { -->
<!--   freq = as.integer(tsp(ts)[3]) -->
<!--   start_year = as.integer(tsp(ts)[1]) -->
<!--   n = length(ts) -->
<!--   trend = NULL -->
<!--   # 1st cursor, controlling pos in whole ts -->
<!--   start_pos = 1 -->
<!--   prev_obj = NULL -->
<!--   while (start_pos <= n) { -->
<!--     x = NULL -->
<!--     y = NULL -->
<!--     # 2nd cursor, controlling accumulating consecutive non-NA -->
<!--     i = start_pos -->
<!--     while (!is.na(ts[i])) { -->
<!--       y = c(y, ts[i]) -->
<!--       x = c(x, i) -->
<!--       i = i + 1 -->
<!--     } -->

<!--     window = n * span -->
<!--     # first time meets NA after consecutive strings of non-NA -->
<!--     if (!is.null(y) && length(y) >= window) { -->
<!--       loess_obj = loess(y ~ x, span=span, degree=deg, family="symmetric", method="loess", control=loess.control(surface="direct", iterations=iter)) -->
<!--       prev_obj = loess_obj -->
<!--       trend = c(trend, loess_obj$fitted) -->
<!--       if (i <= n) { -->
<!--         if (!impute) { -->
<!--           trend = c(trend, NA) -->
<!--         } -->
<!--         else { -->
<!--           pred = predict(object=prev_obj, newdata=data.frame(x=i)) -->
<!--           trend = c(trend, pred) -->
<!--         } -->

<!--       } -->
<!--     } -->
<!--     # some non-NA but length too short -->
<!--     else if (length(y) < window) { -->
<!--       trend = c(trend, rep(NA, length(y))) -->
<!--       if (i <= n) { -->
<!--         trend = c(trend, NA) -->
<!--       } -->
<!--     } -->
<!--     # consecutive NAs -->
<!--     else { -->
<!--       trend = c(trend, NA) -->
<!--     } -->
<!--     start_pos = i + 1 -->
<!--   } -->

<!--   return (ts(trend, start=c(start_year, 1), frequency = freq)) -->
<!-- } -->


<!-- trend.loess = function(ts, deg=2, span=0.75, iter=4, impute=FALSE) { -->
<!--   freq = as.integer(tsp(ts)[3]) -->
<!--   start_year = as.integer(tsp(ts)[1]) -->
<!--   n = length(ts) -->
<!--   trend = NULL -->
<!--   # 1st cursor, controlling pos in whole ts -->
<!--   start_pos = 1 -->
<!--   while (start_pos <= n) { -->
<!--     x = NULL -->
<!--     y = NULL -->
<!--     # 2nd cursor, controlling accumulating consecutive non-NA -->
<!--     i = start_pos -->
<!--     while (!is.na(ts[i])) { -->
<!--       y = c(y, ts[i]) -->
<!--       x = c(x, i) -->
<!--       i = i + 1 -->
<!--     } -->

<!--     window = n * span -->
<!--     # first time meets NA after consecutive strings of non-NA -->
<!--     if (!is.null(y) && length(y) >= window) { -->
<!--       loess_obj = loess(y ~ x, span=span, degree=deg, family="symmetric", method="loess", control=loess.control(surface="direct", iterations=iter)) -->
<!--       trend = c(trend, loess_obj$fitted) -->
<!--       if (i <= n) { -->
<!--         trend = c(trend, NA) -->
<!--       } -->
<!--     } -->
<!--     # some non-NA but length too short -->
<!--     else if (length(y) < window) { -->
<!--       trend = c(trend, rep(NA, length(y))) -->
<!--       if (i <= n) { -->
<!--         trend = c(trend, NA) -->
<!--       } -->
<!--     } -->
<!--     # consecutive NAs -->
<!--     else { -->
<!--       trend = c(trend, NA) -->
<!--     } -->
<!--     start_pos = i + 1 -->
<!--   } -->

<!--   return (ts(trend, start=c(start_year, 1), frequency = freq)) -->
<!-- } -->


<!-- ``` -->



#test
```{r, echo=TRUE, warning=FALSE}
# ts1409_ds = ts1409 - ts(seasonal(ts1409, span=0.5, deg=2), start = c(1980,1), frequency = 365)
start_time = Sys.time()
t2 = trend.loess(ts1409, span=1.5/38)
end_time = Sys.time()
end_time - start_time



plot(ts1409, type="l", col="green", ylim=c(-5,5))
lines(t2, col="red")
lines(stl_obj$time.series[,"trend"], col = "blue")

t3 = trend.loess(ts1409 -  s3, span = 2/38)
plot(ts1409, type="l", col="green", ylim=c(-5,5))
lines(t3, col="red")
lines(stl_obj$time.series[,"trend"], col = "blue")


```















# Convergence test
#==========================================================================================
alternate season & trend extraction difference convergence test

$$X - s_1 = t_1 + \epsilon$$
$$X - t_1 = s_2 + \epsilon$$
$$X - s_2 = t_2 + \epsilon$$
$$\ldots$$
$$X - s_{n-1} = t_{n-1} + \epsilon$$
$$X - t_{n-1} = s_n + \epsilon$$
$$X - S_n = t_n + \epsilon$$


```{r, echo=TRUE}
#iter should be odd
convergence.test = function(ts, iter) {
  ds = NULL
  dt = NULL
  prev_season = rep(0, length(ts))
  prev_trend = rep(0, length(ts))
  for (i in (1:iter)) {
    if (i == 1) {
      curr_season = seasonal.loess(ts, span=1/30)
      curr_trend = trend.loess(ts - curr_season, span=1/38)
      ds = c(ds, mean((curr_season - prev_season)^2, na.rm=TRUE))
      dt = c(dt, mean((curr_trend - prev_trend)^2, na.rm=TRUE))
      prev_season = curr_season
      prev_trend = curr_trend
    }
    else if (i %% 2 != 0) {
      curr_trend = trend.loess(ts - prev_season, span=1/38)
      dt = c(dt, mean((curr_trend - prev_trend)^2, na.rm=TRUE))
      prev_trend = curr_trend
    }
    else {
      curr_season = seasonal.loess(ts - prev_trend, span=1/30)
      ds = c(ds, mean((curr_season - prev_season)^2, na.rm=TRUE))
      prev_season = curr_season
    }
  }
  
  return(data.frame(ds, dt))
}


```



```{r, echo=TRUE, warning=FALSE}
start_time = Sys.time()
convergence_list = convergence.test(ts1409, iter = 31)
end_time = Sys.time()
end_time - start_time
par(mfrow=c(2,1))
plot(convergence_list$dt, col="blue", main="trend")
plot(convergence_list$ds, col="red", main="seasonal")

```



#relative error = sum(r_old^2) - sum(r_new^2)/ sum(r_new^2)

#do while(sum(r_old^2) - sum(r_new^2) < 1 - relative error)

```{r, echo=TRUE}
residual.test = function(ts, iter) {
  relative_errors = NULL

  prev_season = rep(0, length(ts))
  prev_trend = rep(0, length(ts))
  prev_error = rep(0, length(ts))
  for (i in (1:iter)) {
    if (i == 1) {
      curr_season = seasonal.loess(ts, span=1/30)
      curr_trend = trend.loess(ts - curr_season, span=1/38)
      curr_error = ts - curr_season - curr_trend
      
      curr = sum(curr_error^2, na.rm = TRUE)
      prev = sum(prev_error^2, na.rm = TRUE)
      
      relative_errors = c(relative_errors, abs(curr - prev)/curr)
      
      prev_season = curr_season
      prev_trend = curr_trend
      prev_error = curr_error
      
    }
    else if (i %% 2 != 0) {
      curr_trend = trend.loess(ts - prev_season, span=1/38)
      curr_error = ts - curr_trend - prev_season
      curr = sum(curr_error^2, na.rm = TRUE)
      prev = sum(prev_error^2, na.rm = TRUE)
      
      relative_errors = c(relative_errors, abs(curr - prev)/curr)

      prev_trend = curr_trend
      prev_error = curr_error
    }
    else {
      curr_season = seasonal.loess(ts - prev_trend, span=1/30)
      curr_error = ts - curr_season - prev_trend
      curr = sum(curr_error^2, na.rm = TRUE)
      prev = sum(prev_error^2, na.rm = TRUE)
      
      relative_errors = c(relative_errors, abs(curr - prev)/curr)

      prev_season = curr_season
      prev_error = curr_error
    }
  }
  
  return(relative_errors)
}



```



```{r, echo=TRUE, warning=FALSE}
start_time = Sys.time()
residual.test(ts1409, iter = 21)
end_time = Sys.time()
end_time - start_time

```





















# Cross Validation
#==========================================================================================

#cv.seasonal
```{r, echo=TRUE}
cv.seasonal = function(ts, d.max, s.max, fold = 5) {
  start_year = tsp(ts)[1]
  end_year = tsp(ts)[2]
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1 #k = 38 years
  t = ceiling(k/fold) #t = t years/unit
  freq = as.integer(tsp(ts)[3])
  minSMSR = Inf

  best_degree = 0
  best_span = 1*0.05
  
  for (d in (0:d.max)) {
    for (s in (1:(s.max/0.05))) {
      # sum of mean squared error over CVs
      SMSR = 0
      # run 5 times CV
      for (i in (1:fold)) {
        test_index = seq(from=1+(i-1)*t*freq, to=min(i*t*freq, k*freq), by=1)
        test = ts[test_index]
        test = ts(test, start=c(start_year+(i-1)*t, 1), frequency = freq)
        
        train = ts[-test_index]
        train = ts(train, start=c(start_year, 1), frequency = freq)
        seasonality = seasonal.loess(train, deg=d, span=s*0.05)
        
        
        pred = seasonality[1:length(test)]
        residual = (test - pred)
        SMSR = SMSR + mean(residual^2, na.rm = TRUE)
      }
      
      if (SMSR < minSMSR) {
        best_degree = d
        best_span = s*0.05
        minSMSR = SMSR
      }
    }
  }
  
  return (data.frame(best_degree, best_span))
}



```


```{r, echo=TRUE}
start_time = Sys.time()
cv.seasonal(ts1409, d.max=2, s.max=0.5, fold = 2)
end_time = Sys.time()
end_time - start_time

```



#cv.trend
```{r, echo=TRUE}
cv.trend = function(ts, d.max, s.max, fold = 5) {
  start_year = tsp(ts)[1]
  end_year = tsp(ts)[2]
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1 #k = 38 years
  t = ceiling(k/fold) #t = t years/unit
  freq = as.integer(tsp(ts)[3])
  n = length(ts)
  
  
  minSMSR = Inf

  best_degree = 0
  best_span = 1*0.05
  
  for (d in (0:d.max)) {
    for (s in (1:(round(s.max/0.01)))) {
      # sum of mean squared error over CVs
      SMSR = 0
      # run 5 times CV
      remain_index = c(1:n)
      for (i in (1:fold)) {
        test_index = sample(remain_index, size=min(ceiling(n/fold), length(remain_index)), replace=FALSE)
        train_index = c(1:n)[-test_index]

        test = ts[test_index]
        
        train = ts
        train[test_index] = NA
        
        trend = trend.loess(train, deg=d, span=s*0.01)
        pred = as.numeric(predict(trend, newdata=data.frame(x=test_index)))
        residual = (test - pred)
        SMSR = SMSR + mean(residual^2, na.rm = TRUE)
        remain_index = remain_index[-test_index]
      }
      
      if (SMSR < minSMSR) {
        best_degree = d
        best_span = s*0.01
        minSMSR = SMSR
      }
    }
  }
  
  return (data.frame(best_degree, best_span))
}


```


```{r, echo=TRUE}
set.seed(2022)
start_time = Sys.time()
cv.trend(ts1409, d.max=2, s.max=4/38, fold=3)
end_time = Sys.time()
end_time = start_time

```





#========================================================================================
# Model Fitting
<!-- # Classical Decomposition by moving average -->
<!-- ```{r, echo=TRUE} -->
<!-- component = decompose(daily_train, type = "additive", filter = c()) -->
<!-- plot(component) -->
<!-- acf(component$random, na.action = na.pass, lag.max = 365*10) -->
<!-- pacf(component$random, na.action = na.pass, lag.max = 365*5) -->
<!-- pacf(component$random, na.action = na.pass) -->

<!-- ``` -->



# Box-Jenkins Forecast Method Function
```{r, echo=TRUE}
box_jenkins = function(model, ahead) {
  n = length(model$coef)
  mu = model$coef[n]
  preds = NULL
  order = n - 1
  # the list of previous values till today
  queue = window(error, start=c(1991,365-order+1), end=c(1991,365))
  while (ahead != 0) {
    pred = 0
    for (i in (1:order)) {
      pred = pred + as.numeric(queue[order-i+1]) * as.numeric(model$coef[i])
    }
    preds = c(preds, pred)
    queue = c(queue[-1], pred)
    ahead = ahead - 1
  }
  return (ts(preds, start=c(1992,1), frequency = 365))
  
}


```



# ARIMA(5,0,0) model forecast
```{r, echo=TRUE}
lag = 15
future_days = window(daily_test, start=c(1992,1), end=c(1992,lag))


pred1 = predict(model, n.ahead = lag)
predict_values = pred1$pred + loess$time.series[, "trend"][1:lag] + loess$time.series[,"seasonal"][1:lag]
plot(predict_values, col="black", ylim = c(-1.5,-0.5), main="prediction")
lines(future_days, col="red")

#box-jenkins method prediction
pred2 = box_jenkins(model, ahead = lag)
predict_values2 = pred2 + loess$time.series[, "trend"][1:lag] + loess$time.series[,"seasonal"][1:lag]
plot(predict_values2, col="black", ylim = c(-1.5,-0.5), main="prediction")
lines(future_days, col="red")


plot(pred1$pred, col="black", main="prediction")
lines(pred2, col="red")


```






<!-- # package: parallel -->
<!-- ```{r, echo=TRUE} -->
<!-- library(parallel) -->
<!-- library(pbapply) -->

<!-- l <- list(a = c(1,2), b = c(1,2,3))                                 # example list -->
<!-- f <- function(x){return(length(x))}                           # example function -->

<!-- n_cores <- detectCores()                                          # determine number of cores -->
<!-- cl <- makeCluster(n_cores - 1)                                 # use all (except for 1) cores and start cluster -->

<!-- pblapply(l, f, cl = cl)                                                    # use pblapply instead of lapply -->

<!-- stopCluster(cl)                                                            # stop cluster -->



<!-- ``` -->



