---
title: '1409'
author: 'Junsong Tang (64879885)'
date: "xx xx, 2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tseries)
library(zoo)
library(tidyverse)
library(forecast)
load("D:/NSERC/project/processedData/1409_dat.RData")

```

### data pre-processing

# format unit month & unit day

```{r, echo=TRUE}
month = NULL
day = NULL
for (m in (1:12)) {
  if ((m <= 7 && m %% 2 == 1) || (m >= 8 && m %% 2 == 0)){
    month = c(month, rep(m, 31))
    day = c(day, seq(from=1,to=31,by=1))
  } else if (m == 2) {
    month = c(month, rep(m, 28))
    day = c(day, seq(from=1,to=28,by=1))
  } else {
    month = c(month, rep(m, 30))
    day = c(day, seq(from=1,to=30,by=1))
  }
}
months = rep(month, 38)
days = rep(day, 38)

```


#response processing

```{r, echo=TRUE}
y = NULL
years = NULL
for (i in (1:38)) {
  for (j in (1:365)) {
    y = c(y, resp_mat[i,j])
    #N/A & imputation
    ind = (i-1)*365 + j
    if (is.na(y[ind])) {
      y[ind] = y[ind-365]
    }
    years = c(years, 1980+i-1)
  }
}

daily1409 = data.frame(years, months, days, y)
```



# encapsulated function form of daily data process
# input: response(covariate) matrix; boolean(True if impute N/A)
# output: data frame with year, month, day, & response(covariate) values
```{r, echo=TRUE}
dailyFrame = function(data, impute=FALSE) {
  # number of years
  n = length(data[, 1])
  
  # format unit month & days in calender
  month = NULL
  day = NULL
  for (m in (1:12)) {
    if ((m <= 7 && m %% 2 == 1) || (m >= 8 && m %% 2 == 0)) {
      month = c(month, rep(m, 31))
      day = c(day, seq(from = 1, to = 31, by = 1))
    } else if (m == 2) {
      month = c(month, rep(m, 28))
      day = c(day, seq(from = 1,to = 28,by = 1))
    } else {
      month = c(month, rep(m, 30))
      day = c(day, seq(from = 1,to = 30,by = 1))
    }
  }
  # repeat unit format for n years
  months = rep(month, n)
  days = rep(day, n)
  
  
  values = NULL
  years = NULL
  for (i in (1:n)) {
    for (j in (1:365)) {
      values = c(values, data[i,j])
      #N/A & imputation if applicable
      if (impute) {
        ind = (i-1)*365 + j
        if (is.na(values[ind])) {
          values[ind] = values[ind-365]
        }
      }
      years = c(years, 1980+i-1)
    }
  }
  
  return(data.frame(years, months, days, values)) 
}

```



# monthly smooth

```{r, echo=TRUE}
monthly = NULL
yr = NULL
mon = NULL

m = 1
sum = 0
count = 0
ycount = 0
for (i in (1:length(daily1409$months))) {
  if (m == daily1409$months[i] && i < length(daily1409$months)) {
    sum = sum + daily1409$y[i]
    count = count + 1
  } else if (m == daily1409$months[i] && i == length(daily1409$months)) {
    sum = sum + daily1409$y[i]
    count = count + 1 
    avg = sum / count
    monthly = c(monthly, avg)
    mon = c(mon, m)
    yr = c(yr, 1980+ycount)
  } else {
    avg = sum / count
    monthly = c(monthly, avg)
    mon = c(mon, m)
    yr = c(yr, 1980+ycount)
    if (m == 12) {
      ycount = ycount + 1
    }
    m = daily1409$months[i]
    sum = 0 + daily1409$y[i]
    count = 0 + 1
  }
}

monthly1409 = data.frame(yr, mon, monthly)
```





#===============================================================================
### Models

# original series 

```{r, echo=TRUE}
res_ts=ts(monthly1409$monthly, start = c(1980,1), frequency = 12)
plot(res_ts)
acf(res_ts, lag.max = 12*5, main = "acf of monthly response")
pacf(res_ts, lag.max = 12*5, main = "pacf of monthly respoonse")
# no obvious up or down trend, mean is non zero, but obvious seasonality
```


#(1) SARIMA(0,0,0)X(0,1,0)

```{r, echo=TRUE}
sarima.000.010 = arima(res_ts, order=c(0, 0, 0), seasonal=list(order=c(0, 1, 0), period=12), include.mean = TRUE)
plot(sarima.000.010$residuals,main="SARIMA (0, 0, 0) x (0, 1, 0)", ylab="Residuals")
acf(sarima.000.010$residuals,lag.max=12*10, main="Autocorrelation")
pacf(sarima.000.010$residuals,lag.max=12*10, main="Partial Autocorrelation")
# acf shows gradually decaying pattern at seasonal lags
# pacf cuts off at seasonal lag 4(or lag 5? lag 6?)
# need seasonal AR component
```


#(2) add AR(P=5) component

```{r, echo=TRUE}
sarima.000.510 = arima(res_ts, order=c(0, 0, 0), seasonal=list(order=c(5, 1, 0), period=12), include.mean = TRUE)
plot(sarima.000.510$residuals,main="SARIMA (0, 0, 0) x (5, 1, 0)", ylab="Residuals")
acf(sarima.000.510$residuals,lag.max=12*10, main="Autocorrelation")
pacf(sarima.000.510$residuals,lag.max=12*10, main="Partial Autocorrelation")
# pacf cuts off at lag 1, 
# acf cuts off at lag 4, need non seasonal MA component
```


#(3) add MA(q=4) component
```{r, echo=TRUE}
sarima.004.510 = arima(res_ts, order=c(0, 0, 4), seasonal=list(order=c(5, 1, 0), period=12), include.mean = TRUE)
plot(sarima.004.510$residuals,main="SARIMA (0, 0, 4) x (5, 1, 0)", ylab="Residuals")
acf(sarima.004.510$residuals,lag.max=12*10, main="Autocorrelation")
pacf(sarima.004.510$residuals,lag.max=12*10, main="Partial Autocorrelation")
# acf cuts of at lag 0
# pacf  cuts off lag 1
# seems ok

```


# diagnostics
```{r, echo=TRUE}
tsdiag(sarima.004.510)
# Ljung Box stats are good
sarima.004.510$aic
# AIC = 1201.486, lowest from all models fitted
qqnorm(sarima.004.510$residuals, ylab = "residue quantile")
qqline(sarima.004.510$residuals)
# heavy tail for residuals
sarima.004.510$sigma2
```





# log response(log monthly) and fitting
```{r, echo=TRUE}
log_ts = log(res_ts)

model1 = arima(log_ts, order=c(0, 0, 0), seasonal=list(order=c(0, 1, 0), period=12), include.mean = TRUE)
acf(model1$residuals,lag.max=12*10, main="Autocorrelation")
pacf(model1$residuals,lag.max=12*10, main="Partial Autocorrelation")


# add AR(P=5) seasonal
model2 = arima(log_ts, order=c(0, 0, 0), seasonal=list(order=c(5, 1, 0), period=12), include.mean = TRUE)
acf(model2$residuals,lag.max=12*10, main="Autocorrelation")
pacf(model2$residuals,lag.max=12*10, main="Partial Autocorrelation")


# add MA(q=5) non-seasonal
model3 = arima(log_ts, order=c(0, 0, 5), seasonal=list(order=c(5, 1, 0), period=12), include.mean = TRUE)
plot(model3$residuals, main="SARIMA(0,0,5) X (5,1,0)")
acf(model3$residuals,lag.max=12*10, main="Autocorrelation")
pacf(model3$residuals,lag.max=12*10, main="Partial Autocorrelation")

tsdiag(model3)
qqnorm(model3$residuals, ylab = "residue quantile")
qqline(model3$residuals)
```



# train & test
```{r, echo=TRUE}
# partition train set & test set(10% total size)
train = window(log_ts, start=c(1980,1), end=c(2013,12))
test = window(log_ts, start=c(2014,1), end=c(2017,12))

model = arima(train, order=c(0, 0, 5), seasonal=list(order=c(5, 1, 0), period=12), include.mean = TRUE)
acf(model$residuals,lag.max=12*10, main="Autocorrelation")
pacf(model$residuals,lag.max=12*10, main="Partial Autocorrelation")

# usual prediction
pred1 = predict(model, n.ahead = 12*4)
lb = pred1$pred - qnorm(0.975) * pred1$se
ub = pred1$pred + qnorm(0.975) * pred1$se
plot(pred1$pred, main = "Regular Predictions")
lines(test, col="red")
# lines(lb, col = "blue", lty = 2)
# lines(ub, col = "blue", lty = 2)
legend("bottomleft",legend=c("predict value","actual value"), lty=1, col=c("black","red"))


RMSE_sarima = sqrt(mean((pred1$pred-test)^2))
```


# holt-winters method
```{r, echo=TRUE}
model_hw = HoltWinters(train)
pred2 = predict(model_hw, n.ahead = 12*4)
plot(pred2, main="HoltWinters Predictions")
lines(test, col="red")
legend("bottomleft",legend=c("predict value","actual value"), lty=1, col=c("black","red"))

RMSE_HW = sqrt(mean((pred2-test)^2))

```




# Decomposition method & fitting
```{r, echo=TRUE}
# usual decomposition
# component = decompose(train, type = "additive")
# plot(component)
# error = component$random
# error = na.remove(error)
# #correlogram
# acf(error, lag.max = 12*5)
# pacf(error, lag.max = 12*5)
# #qq plot
# qqnorm(error)
# qqline(error)



# Loess decomposition
train.stl = stl(train, s.window = "periodic")
plot(train.stl)
# check residuals
error = train.stl$time.series[,"remainder"]
qqnorm(error)
qqline(error)


#de-seasonalize
train.ds = train - train.stl$time.series[,"seasonal"]
plot(train.ds, main ="de-seasonalized")
acf(train.ds, main = "acf ds")
pacf(train.ds, main = "pacf ds")

# re-center
mu = mean(train.ds)
train.rc = train.ds - mu
plot(train.rc, main ="de-seasonalized & re-centered ts")
acf(train.rc, lag.max = 12*5, main = "acf ds + rc")
pacf(train.rc, lag.max = 12*5, main = "pacf ds + rc")


# acf decaying gradually and pacf cuts off at lag 2, so we fit AR(2)
model.decomp = arima(train.rc, order=c(2,0,0))
tsdiag(model.decomp)


# prediction
train.pred = predict(model.decomp, n.ahead = 12*4)

train.pred$pred = train.pred$pred + mu + train.stl$time.series[,"seasonal"][1:48]

# prediction intervals
lb = train.pred$pred - qnorm(0.975)*train.pred$se
ub = train.pred$pred + qnorm(0.975)*train.pred$se


plot(train.pred$pred, main = "Predictions")
lines(test, col="red")
# lines(lb, col = "blue", lty = 2)
# lines(ub, col = "blue", lty = 2)
legend("bottomleft",legend=c("predict value","actual value"), lty=1, col=c("black","red"))

RMSE_Decomp = sqrt(mean((train.pred$pred-test)^2))

```

## Comments: 
RMSE for step-by-step sarima: 0.463787;
RMSE for Holt_winter: 0.4377578;
RMSE for decompositon fit: 0.4422446;

































#===============================================================================
# Daily unimputed data
```{r, echo=TRUE}
daily = dailyFrame(resp_mat)

# identifying the range of NA
daily[is.na(daily$values),]

# subset of log-daily from 1980.1 - 1994.12(15 years)
daily_ts = ts(log(daily$values), start=c(1980,1), end = c(1994,12), frequency = 365)
plot(daily_ts)
acf(daily_ts, lag.max = 365*5, main = "acf of daily response")
pacf(daily_ts, lag.max = 365*5, main = "pacf of daily respoonse")

# train set 12 years, test set 3 years
daily_train = window(daily_ts, start=c(1980,1), end=c(1991,12))
daily_test = window(daily_ts, start=c(1992,1), end=c(1994,12))


# 

```



# Loess decomposition on daily
```{r, echo=TRUE}
daily.train.stl = stl(daily_train, s.window = "periodic")
plot(daily.train.stl)
# check residuals
error_daily = daily.train.stl$time.series[,"remainder"]
qqnorm(error_daily)
qqline(error_daily)


#de-seasonalize
daily.train.ds = daily_train - daily.train.stl$time.series[,"seasonal"]
plot(daily.train.ds, main ="de-seasonalized")
acf(daily.train.ds, lag.max = 365*5, main = "acf ds")
pacf(daily.train.ds, lag.max = 365*5, main = "pacf ds")

# re-center
daily.mu = mean(daily.train.ds)
daily.train.rc = daily.train.ds - daily.mu
plot(daily.train.rc, main ="de-seasonalized & re-centered ts")
acf(daily.train.rc, lag.max = 365*5, main = "acf ds + rc")
pacf(daily.train.rc, lag.max = 365*5, main = "pacf ds + rc")


```



# Fourier method to decompose long seasonality(365)
```{r, echo=TRUE}

daily_model = auto.arima(daily_train, xreg = fourier(daily_train, K=2))
# seasonality = fourier(daily_train, K = 2)
# Nt = daily_train-seasonality
tsdiag(daily_model)
qqnorm(daily_model$residuals)
qqline(daily_model$residuals)


```


```{r, echo=TRUE}
# prediction


plot(forecast(daily_model, h=3*365, xreg=fourier(daily_train, K=2, h=3*365)))
lines(daily_test, col="red")

# prediction intervals
# lb = daily.pred$pred - qnorm(0.975)*daily.pred$se
# ub = daily.pred$pred + qnorm(0.975)*daily.pred$se


# plot(daily.pred$pred, main = "Predictions")
# lines(daily_test, col="red")
# # lines(lb, col = "blue", lty = 2)
# # lines(ub, col = "blue", lty = 2)
# legend("bottomleft",legend=c("predict value","actual value"), lty=1, col=c("black","red"))

```

