---
title: "simulation"
author: "Junsong Tang"
date: '2022'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(tseries)
library(stlplus)
library(forecast)
stlp1409 = stlplus(ts1409, n.p=365, s.window="periodic", inner=2, outer=10)
```


# tool functionalities...
```{r}
mse = function(X, Y) {
  if (length(X) != length(Y)) {
    print("lengths differ!")
    return ()
  }
  else {
    return (mean((X - Y)^2, na.rm = TRUE))
  }
}



# n order difference in vector/time series
difference = function(X, order) {
  if (order == 1) {
    result = NULL
    n = length(X)
    for (i in (2 : n)) {
      result = c(result, (X[i]-X[i-1]))
    }
    return (result)
  }
  else {
    result = difference(X,1)
    return (difference(result, (order-1)))
  }
}


ma = function(Xt, window) {
  n = length(Xt)
  # window >= 2
  if (window %% 2 == 0 & window > 0) {
    window = window - 1
  }
  r = (window - 1) / 2
  Yt = NULL

  
  for (i in (1 : n)) {
    if (i <= r || i > (n-r)) {
      avg = NA
    }
    else {
      avg = mean(Xt[(i-r) : (i+r)])
    }
    Yt = c(Yt, avg)
  }
  
  return (Yt)
}

monotone = function(X) {
  n = length(X)
  if (n == 1) {
    monotone = TRUE
  }
  else {
    i = 2
    monotone = X[i] <= X[i-1]
    while (i <= n && monotone) {
      i = i + 1
      if (i <= n) {
        monotone = X[i] <= X[i-1]
      }
      else {
        monotone = TRUE
      }
      
    }
  }
  return (monotone)

}



partialsums = function(X) {
  n = length(X)
  sum = 0
  par_sums = NULL
  for (i in (1:n)) {
    curr = sum + X[i]
    par_sums = c(par_sums, curr)
    sum = sum + X[i]
  }
  return (par_sums)
}

```



#simulation

<!-- #==================================================================================== -->
<!-- # sparse data -->
<!-- ```{r, echo=TRUE} -->
<!-- # period = 10 -->
<!-- # 10 periods -->
<!-- t = c(1 : 100) -->
<!-- w = (2*pi)/10 -->
<!-- sea = 5*sin(w*t) -->
<!-- tr = 0.0001 * (t - 20)* (t - 50)* (t - 100) -->
<!-- # err = rt(length(t), df = 5) #any fat tail distr and picking one side tail -->
<!-- err = rnorm(length(t)) -->
<!-- # corr = arima.sim(list(order = c(), ar = 0.5, ma = 0.3), n = length(t)) -->
<!-- Xt = sea + tr + err -->
<!-- Xt = ts(Xt, start=c(2022,1), frequency = 10) -->

<!-- ``` -->




<!-- ```{r, echo=TRUE, warning=FALSE} -->
<!-- # d = decompose(Xt, s.window = c(1,20), t.window = 40, iter =3) -->
<!-- dstlp = stlplus::stlplus(Xt, n.p=10, s.window="periodic", inner=2, outer=10) -->

<!-- # t = trend(Xt, window = 95) -->
<!-- # s = seasonal(Xt - t, window = 30) -->
<!-- # t_ma = trend.ma(Xt) -->
<!-- ``` -->


<!-- ```{r, echo=TRUE} -->
<!-- best_tw = function(tr) { -->
<!--   min_MSE = Inf -->
<!--   best_w = 10 -->
<!--   for (w in seq(from=10, to=100, by=1)) { -->
<!--     t = trend(Xt, window = w) -->
<!--     MSE = mean(sqrt((t - tr)^2)) -->
<!--     if (MSE < min_MSE) { -->
<!--       min_MSE = MSE -->
<!--       best_w = w -->

<!--     } -->

<!--   } -->

<!--   return (best_w) -->
<!-- } -->


<!-- ``` -->



<!-- ```{r, echo=TRUE, warning=FALSE} -->
<!-- t = trend(Xt, window = best_tw(tr)) -->
<!-- res = cv.seasonal(Xt - t, window.min=10, window.max=50) -->
<!-- s = seasonal(Xt - t, deg=res$best_deg, window=res$best_window) -->


<!-- ``` -->























#==================================================================================
# rich data

# setup
```{r, echo = TRUE}
ndays <- 365; nyears <- 10; start_time <- c(2012,1)
n <- ndays * nyears
t <- 1 : (n)
w <- (2*pi)/ndays 


sea <- 5*sin(w*t) #seasonal

# heavy trend
tr <- 0.1*(t/ndays - 1)* (t/ndays- 6)* (t/ndays - 10)

# flat trend

tr2 = stlp1409$data[,"trend"][1:3650]


# # linear trend
# tr3 = 0.001 * t + 1

```


# errors
```{r, echo=TRUE}
c = 1


# White noise
# err = c * rnorm(length(t))



# ARMA error
err = c * arima.sim(n=ndays*nyears, list(ar = c(0.5), ma = c(1, 0.5)))


# AR error
# err = c * arima.sim(n=ndays*nyears, list(ar = c(0.8, -0.5)))


# MA error
# err = c * arima.sim(n=ndays*nyears, list(ma = c(1, 0.5)))
```




# adding outliers
```{r, echo=TRUE}
# adding outliers
set.seed(12)
prop = 0.01
idx <- sample(n, round(prop*n))


# with mean = 30 , sd = 2
# Xt[idx] <- rnorm(length(idx), 30, 2)
# 
# with mean = 0 , sd = 20
# Xt[idx] <- rnorm(length(idx), 0, 10)

# err[idx] <- rnorm(length(idx), 30, 2)

err[idx] <- rnorm(length(idx), 0, 3)

# Xt <- ts(Xt, start=start_time , frequency = ndays)
# plot(Xt, type = "p")


#tmp <- rbinom(length(idx), 1,0.5)
#Xt[idx] <- tmp*rnorm(length(idx), 25, 2) + (1-tmp)*rnorm(length(idx), -25, 2)

```






# forming Xt(ts) (later try different seeds for at least 50 times!!)
```{r}
Xt = tr + err


Xt <- ts(Xt, start=start_time , frequency = ndays)
plot(Xt, col = "grey", type = "b", main = "Xt")
lines(ts(tr, start =start_time, frequency = ndays), col = "red")
legend("topright",legend=c("true trend"), lty=1,
       col=c("red"))


```







```{r}
Yt = ts(c(Xt[1], difference(Xt, order = 1)), start=start_time , frequency = ndays)
acf_Y = acf(Yt, lag.max = n)
acf_S = acf(difference(err, order = 1), lag.max = n)


```


```{r}
for (span in seq(from = 1, to = 0.1, by = -0.1)) {

  t_i = trend(Yt, deg = 2, window = span * n)
  r_i = Yt - t_i
  
  
  dt = ts(c(tr[1], difference(tr, order=1)), start=start_time , frequency = ndays)
    
  par(mfrow=c(1,1))
  # raw
  plot(Yt, type = "p", col="grey", main = paste0("span = ", span))
  
  # fit
  lines(t_i, lwd = 2, type = "l", col = "blue")
  # true trend & seasonal
  lines(ts(difference(tr, order=1), start =start_time, frequency = ndays), lwd = 2, type = "l", col="red")
  legend("topright",legend=c("trend fit", "true trend"), lty=1,
       col=c("blue", "red"))
  print(mse(t_i, dt))
}



```





```{r}
tt = trend(Yt, deg = 1, window = 1 * n)
plot(tt, type = "b", col = "red", ylim = c(-0.1, 0.1))
lines(dt, type = "b", col = "blue")
fit_tr = ts(partialsums(tt), start=start_time , frequency = ndays)
plot(fit_tr, type = "b", col = "red")
lines(ts(tr, start =start_time, frequency = ndays), type = "b", col = "blue")


```











```{r}
acf_R = acf(err, lag.max = n)
sum(abs(as.vector(acf_Y$acf)) <= (1.96/sqrt(n))) / length(as.vector(acf_Y$acf))
sum(abs(as.vector(acf_R$acf)) <= (1.96/sqrt(n))) / length(as.vector(acf_R$acf))

# for (span in seq(from = 1, to = 0.1, by = -0.1)) {
#   t_i = trend(Xt, deg = 2, window = span * n)
#   r_i = Xt - t_i
#   acf_r = acf(r_i, lag.max = n, plot = FALSE)
#   # m = length(acf_r$acf)
#   within = abs(acf_r$acf[(l+1):n]) <= (1.96/sqrt(n))
#   prop = sum(within) / length(within[(l+1):n])
# 
#   if (prop >= max_prop) {
#     best_span = span
#     max_prop = prop
#   }
# }
# max_prop
# best_span


```

<!-- #stl-plus result -->
<!-- ```{r, echo=TRUE} -->
<!-- ## stl_plus -->
<!-- res_stl_plus <- stlplus::stlplus(Xt, n.p=ndays, s.window="periodic", inner=2, outer=10) -->

<!-- ## stl does not work "series is not periodic or has less than two periods"; check frequency first ...  -->
<!-- res_stl <- stl(Xt,  s.window="per", inner=2, outer=10) -->

<!-- ## check results:  -->
<!-- plot(Xt, type = "p", col="grey") -->


<!-- # lines(ts(res_stl_plus$data[,"seasonal"], start =start_time , frequency = ndays), lwd = 2, type = "l", col = "blue") -->
<!-- lines(ts(res_stl_plus$data[,"trend"], start = start_time , frequency = ndays), lwd = 2, type = "l", col = "blue") -->


<!-- # true trend & seasonal -->
<!-- # lines(ts(sea, start =start_time, frequency = ndays), lwd = 2, type = "l", col="red") -->
<!-- lines(ts(tr2, start =start_time, frequency = ndays), lwd = 2, type = "l", col="red") -->


<!-- legend("topright",legend=c("trend fitted", "true trend"), lty=1, -->
<!--        col=c("blue", "red")) -->
<!-- ``` -->







#acf of simulated data
```{r}
acf_X = acf(Xt, lag.max = 100, main = "acf of trend + remainder")
acf_R = acf(err, lag.max = 100, main = "acf of true remainder")
# acf_T = acf(tr2, lag.max = 3650, main = "acf of true trend")

acov_X = acf(Xt, type = "covariance",  lag.max = n, main = "acf of trend + remainder")
acov_R = acf(err, type = "covariance", lag.max = n, main = "acf of true remainder")

acf = as.vector(acf_X$acf)

# plot(log(acf_R$acf - min(acf_R$acf)), type = "b", col = "red")
# lines(log(acf_X$acf), type = "b", col = "blue")


# plot(acov_X$acf - 7, type = "b", col = "blue")
# lines(acov_R$acf, type = "b", col = "red")

plot(difference(as.vector(acov_R$acf), order = 1)[1:100], type = "b", col = "blue")
lines(difference(as.vector(acov_X$acf), order = 1)[1:100], type = "b", col = "red")


plot(abs(difference(as.vector(acf_R$acf), order = 1))[1:50], type = "b", col = "blue")
lines(abs(difference(as.vector(acf_X$acf), order = 1))[1:50], type = "b", col = "black")
lines(y=rep(2*1.96/sqrt(n), 1000), x=c(1:1000), col = "red")
lines(y=rep(1.96/sqrt(n), 1000), x=c(1:1000), col = "red")
legend("topright",legend=c("error", "Xt"), lty=1,
       col=c("blue", "black"))
```




<!-- ```{r} -->
<!-- acf_1409 = acf(ts1409-seasonal(ts1409, window = 60 * 38), na.action = na.exclude, lag.max = 100) -->

<!-- acov_1409 = acf(ts1409-seasonal(ts1409, window = 60 * 38), type = "covariance", na.action = na.exclude, lag.max = 100) -->

<!-- ``` -->


# method 1
determine the critical lag: $$\hat{h}$$ such that $$\rho(\hat{h}+1) > \rho(\hat{h})$$, then $$\forall h \geq \hat{h}$$, calculate # (acf < bound) / (n - h) when searching span, stop at the span when this proportion is maximized. Also, we want the MSE of fitted the trend to be as small as possible.



```{r}
l = 1

while (acf[l+1] <= acf[l] || abs(difference(acf, order=1))[l] >= (2*1.96/sqrt(n))) {
  l = l + 1
}

l

```


<!-- ```{r} -->
<!-- span = 0.2 -->
<!-- t_i = trend(Xt, deg = 2, window = span * n) -->
<!-- r_i = Xt - t_i -->
<!-- acf_r = acf(r_i, lag.max = n, plot = FALSE) -->
<!-- # m = length(acf_r$acf) -->
<!-- within = abs(acf_r$acf[(l+1):n]) <= (1.96/sqrt(n)) -->
<!-- sum(within) / length(within[(l+1):n]) -->
<!-- plot(acf_r) -->
<!-- # while (span >= 0.2 && sum(within) / length(within) < 0.95 ) { -->
<!-- #   span = span - 0.1 -->
<!-- #   t_i = trend(Xt, deg = 2, window = span * n) -->
<!-- #   r_i = Xt - t_i -->
<!-- #   acf_r = acf(r_i, lag.max = n, plot = FALSE) -->
<!-- #   m = length(acf_r$acf) -->
<!-- #   within = abs(acf_r$acf[(l+1):m]) <= (1.96/sqrt(n)) -->
<!-- # } -->

<!-- #maximize the proportion of #(acf<bound) / #acf -->


<!-- ``` -->



```{r}
# max_prop = 0
best_span = 1
min_d = Inf

for (span in seq(from = 1, to = 0.1, by = -0.1)) {
  t_i = trend(Xt, deg = 2, window = span * n)
  r_i = Xt - t_i
  acf_r = acf(r_i, lag.max = n, plot = FALSE)
  # m = length(acf_r$acf)
  ref_prop = sum(abs(as.vector(acf_Y$acf)) <= (1.96/sqrt(n))) / length(as.vector(acf_Y$acf))
  prop = sum(abs(as.vector(acf_r$acf)) <= (1.96/sqrt(n))) / length(as.vector(acf_r$acf))
  # within = abs(acf_r$acf[(l+1):n]) <= (1.96/sqrt(n))
  # prop = sum(within) / length(within[(l+1):n])
  d = abs(prop - ref_prop)
  print(d)

  if (d <= min_d) {
    best_span = span
    min_d = d
  }
  

  # if (prop >= max_prop) {
  #   best_span = span
  #   max_prop = prop
  # }
}

best_span


```


# method 2
Assume $$X_t = m_t + Y_t$$
# acov(Xt) and acov(Yt) has constant shift at small lag h compared to large n, 

```{r}
plot(as.vector(acov_X$acf)[1:100], type = "h", col = "blue", ylim = c(-1,10))
lines(as.vector(acov_R$acf)[1:100], type = "h", col = "red")



```



```{r}
min_v = Inf
best_span = 1

for (span in seq(from = 1, to = 0.1, by = -0.1)) {
  t_i = trend(Xt, deg = 2, window = span * n)
  r_i = Xt - t_i
  acov_r = acf(r_i, type = "covariance", lag.max = n, plot = FALSE)
  acf_r = acf(r_i, lag.max = n, plot = FALSE)
  D = difference(as.vector(acov_X$acf)[1:(l+1)], order=1)
  d = difference(as.vector(acov_r$acf)[1:(l+1)], order=1)
  # if (v <= min_v) {
  #   best_span = span
  #   min_v = v
  # }
  plot(as.vector(acov_X$acf)[1:100], type = "h", col = "blue", ylim = c(-1,10))
  lines(as.vector(acov_r$acf)[1:100], type = "h", col = "red")
  print(abs(d))
}
# min_v
# best_span

```

# method 3
# find new critical lag: h such that we encounter second time for the increase in acf after the second monotone decrease in acf

```{r}
acf(difference(Xt, order=1), type = "covariance")
plot(acov_R$acf[1:100], type = "h")
```




#cheat check
```{r}
shift = var(tr)
plot(as.vector(acov_X$acf)[1:100] - shift, type = "b", col = "red")
lines(as.vector(acov_R$acf)[1:100], type = "b", col = "blue")

```




# different span result comparison
```{r}
# for (span in seq(from = 1, to = 0.1, by = -0.1)) {
#   init_season = rough.season(Xt)
#   offset = mean(init_season)
#   init_season = init_season - offset
#   t_i = trend(Xt-init_season, deg = 2, window = span * n)
#   opt = cv.seasonal(Xt - t_i, span.min = 0.01, span.max = 0.2)
#   s_i = seasonal(Xt - t_i, deg = opt$best_deg, window = opt$best_span * length(Xt))
#   r_i = Xt - t_i - s_i
#   par(mfrow=c(1,1))
#   # raw
#   plot(Xt, type = "p", col="grey")
#   
#   # fit
#   lines(t_i, lwd = 2, type = "l", col = "blue")
#   lines(s_i, lwd = 2, type = "l", col = "blue")
#   
#   # true trend & seasonal
#   lines(ts(sea, start =start_time, frequency = ndays), lwd = 2, type = "l", col="red")
#   lines(ts(tr, start =start_time, frequency = ndays), lwd = 2, type = "l", col="red")
#   
#   # spec = spectrum(Xt - s_i, na.action=na.exclude)
#   # spec_chart = data.frame(period=1/t_i$freq, frequency=t_i$freq, spectrum=t_i$spec)
#   # head(spec_chart[order(spec_chart$spectrum, decreasing=TRUE),], 10)
#   par(mfrow = c(2,1))
#   plot(t_i)
#   plot(r_i)
#   
# }

# plot(as.vector(acf_Y$acf - acf_S$acf), type = "b")

# no seasonal
for (span in seq(from = 1, to = 0.1, by = -0.1)) {

  t_i = trend(Xt, deg = 2, window = span * n)
  r_i = Xt - t_i
  
  dr_i = difference(r_i, order = 1)
  
  acf_s = acf(dr_i, lag.max=n, plot = FALSE)
  
    
  par(mfrow=c(1,1))
  # raw
  plot(Xt, type = "p", col="grey", main = paste0("span = ", span))
  
  # fit
  lines(t_i, lwd = 2, type = "l", col = "blue")
  # true trend & seasonal
  lines(ts(tr, start =start_time, frequency = ndays), lwd = 2, type = "l", col="red")
  legend("topright",legend=c("trend fit", "true trend"), lty=1,
       col=c("blue", "red"))
  # spec = spectrum(Xt - s_i, na.action=na.exclude)
  # spec_chart = data.frame(period=1/t_i$freq, frequency=t_i$freq, spectrum=t_i$spec)
  # head(spec_chart[order(spec_chart$spectrum, decreasing=TRUE),], 10)
  
  # par(mfrow = c(2,1))
  # plot(t_i)
  # plot(r_i)
  
  # plot(as.vector(acf_Y$acf - acf_s$acf), type = "b")
  
  # par(mfrow = c(1,2))
  # acf(dr_i,lag.max = 20, ylim = c(-1, 2))
  # acf(Yt, lag.max = 20, ylim = c(-1, 2))
  
  # print(monotone(as.vector(acf(t_i, lag.max = 50, plot=FALSE)$acf)))
  # print(sum((as.vector(acf_Y$acf) - as.vector(acf_s$acf))^2))
  
  # d_r = abs(difference(as.vector(acf_r$acf), order = 1))
  # d_R = abs(difference(as.vector(acf_R$acf), order = 1))
  # par(mfrow=c(1,1))
  # plot(d_r, type = "b", col = "blue",ylim = c(0,1))
  # lines(d_R, type = "b", col = "red")
  # legend("topright",legend=c("fitted error", "true error"), lty=1,
  #      col=c("blue", "red"))
  
  
}


```



```{r, echo=TRUE}
plot(Xt, type = "b", col="grey")

lines(ts(dstlp$data[,"seasonal"], start =c(2022,1), frequency = 10), type = "l", col = "blue")
lines(ts(dstlp$data[,"trend"], start =c(2022,1), frequency = 10), type = "l", col = "blue")
# # lines(ts(dstlp$data[,"seasonal"]+dstlp$data[,"trend"], start =c(2022,1), frequency = 10), type = "l", col = "blue")
lines(s, type = "l", col = "red")
lines(t, type = "l", col = "red")
lines(ts(sea, start =c(2022,1), frequency = 10), type = "l", col="green")
lines(ts(tr2, start =c(2022,1), frequency = 10), type = "l", col="green")
# legend("bottomright",legend=c("true","loess", "stlplus"), lty=1,
#        col=c("green","red", "blue"))

```



```{r}
acf(ts1409, na.action = na.exclude, lag.max = length(ts1409))
# acf(stlp1409$data[,"trend"] + stlp1409$data[,"remainder"], na.action = na.exclude, lag.max = 1000)
# acf(ts1409 - seasonal(ts1409, window = 60*38), na.action = na.exclude, lag.max = 1000)

acf(difference(ts1409, order=1), na.action = na.exclude, lag.max = 200)
```














#======================================================================================
# jasmine's code
```{r include = TRUE}



#acf_original <- acf(as.numeric(dat_obj$Xt))
acf_original <- acf(as.numeric(Xt))  ## if we know that the error looks like 
#par(mfrow = c(1,1))
#acf(as.numeric(Xt))

bws <- seq(0.1,1,0.05)
acf_select <- rep(NA, length(bws))
i <- 1; type <- "K-fold"
par(mfrow = c(1,2))
for(bw in bws){
  trend_obj <- extract.trend(Xt,  type  = type, trend_bw = bw)
  x_trend <- trend_obj$x_trend
  x_remainder <- Xt - x_trend 
  plot(ts(x_trend, start = start_time, frequency = ndays), type = "l",main = bw)
  lines(ts(tr2, start = start_time, frequency = ndays),  col = "red")
  acf_obj <- acf(as.numeric(x_remainder))
  acf_select[i] <- sum( ((as.numeric(unlist(acf_obj)))[1:4] -(as.numeric(unlist(acf_original)))[1:4])^2)
  i <- i+1
}
```


```{r, echo=TRUE}
## try to mimic what stl_plus is doing ... 
cycle_indices <- rep(rep(1:ndays), nyears)
cs1 <- cs2 <- rep(1:ndays) # head and tail 
n.p <- ndays
c_tilde <- rep(NA, n)
for(i in 1:ndays){
  idx <- which(cycle_indices== i)
  c_tilde[idx] <- mean(Xt[which(cycle_indices==i)])
}
plot(c_tilde)
ma <- function(x, n = 5){filter(x, rep(1 / n, n), sides = 2)}
c_tilde_ma <- ma(c_tilde, n.p) # this is a constant 
L <- c_tilde_ma
seasonal <- c_tilde - L   # not smooth; just the average of days per year   
D <- as.numeric(Xt) - seasonal # this is trend 

# smooth the trend to ignore things < 20 per days .... 
# Let's try to code Jason's method 

init.trend <- function(x, ndays, nyears){
  
  x_sub <- matrix(x, ncol = ndays)
  seasonal_med <- apply(x_sub, 2, median)
  #trend_rough <- ma(x-rep(seasonal_med, nyears), ndays)
  tmp <- as.numeric(x-rep(seasonal_med, nyears))
  time_pt <- 1:length(x)  
  trend_init <- loess(tmp~time_pt, family = "symmetric") # might be a little oversmoothed... so seasonal may have short term, but should be 
  #okay since the short term differ in years, and likely not be picked up by the next seasonal smoothing 
  return(predict(trend_init, time_pt))
}


cv.seasonal <- function(x_sub, candidate_bw){  
  
  time_pt <- 1:ncol(x_sub)  
  cv_err <- 0
  for(i in 1:nrow(x_sub)){
    tmp <- as.numeric(x_sub[-i,])
    loess_obj <- loess(tmp~rep(time_pt, nrow(x_sub)-1), span =  candidate_bw,  family = "symmetric")
    
    cv_err <- cv_err + sum( (x_sub[i,] - predict(loess_obj, time_pt))^2)
  }
  cv_err  <- cv_err/nrow(x_sub) # average prediction error per year	
  return(cv_err)
}

fit.seasonal <- function(x_sub, bw){
  
  time_pt <- 1:ncol(x_sub)  
  tmp <- as.numeric(x_sub)
  loess_obj <- loess(tmp~rep(time_pt, nrow(x_sub)), span =  bw,  family = "symmetric") 
  fitted <- predict(loess_obj, time_pt)

}

# if bw too large, trend flat; else trend rough 
decompose_trend_remainder <- function(x_trend_remainder, bw){
  
  time_pt <- 1:length(x_trend_remainder)  
  tmp <- as.numeric(x_trend_remainder)
  loess_obj <- loess(tmp~time_pt, span =  bw,  family = "symmetric") 
  fitted_trend <- predict(loess_obj, time_pt)
  return(list(fitted_trend = fitted_trend, fitted_remainder = x_trend_remainder - fitted_trend))
  
}


perform.decompose  <- function(x, ndays, nyears, niter){
  
  x_detrend <- x - init.trend(x, ndays, nyears) # first detrend
  
  for(i in 1:niter){
    
    x_sub <- matrix(x_detrend, ncol = ndays)
    opt_bw <- optim(par = 0.5, fn = function(candidate_bw){cv.seasonal(x_sub, candidate_bw)}, lower = 0.01, upper = 1, method = "Brent") 
    x_seasonal <- fit.seasonal(x_sub, opt_bw$par)
    x_seasonal_rep <- rep(x_seasonal, nyears)
    x_trend_remainder <- x - x_seasonal_rep
    
    par(mfrow = c(3,2))
    
    par(mfrow = c(1,1))
    
    for(bw in c(0.1, 0.5, 2)){
      
      tmp <- decompose_trend_remainder(x_trend_remainder, bw)
      plot(tmp$fitted_trend, main = var(tmp$fitted_trend), ylab = "fitted trend")
      plot(tmp$fitted_remainder, main = var(tmp$fitted_remainder), ylab = "fitted residual")
      
    }
    
    
    #library("funtimes")
    time_points <- 1:(ndays* nyears) 
    tmpp <- loess(tmp$fitted_remainder~time_points,  family = "symmetric")   # not super smooth but should be good enough .. 
    tmpp <- predict(tmpp, time_points)
    plot(tmp$fitted_trend)
    plot(tmpp)
    
    plot(abs(tmp$fitted_trend))
    plot(abs(tmpp))
    
    
    aa <- quantile(abs(tmp$fitted_trend), seq(0,1,0.1))
    bb <- quantile(abs(tmpp), seq(0,1,0.1))
    plot(bb~aa)
    abline(a = 0, b = 1, col = "red")
    
    
    t.test(abs(tmpp))
    
    tmpp <- tmpp[1:200]
    time_points1 <- 1:length(tmpp)
    
    wavk_test(tmpp ~ time_points1, factor.length = "adaptive.selection")
    
    tmpp  <- 0.5*time_points1 + rnorm(length(tmpp))
    wavk_test(tmpp ~ time_points1 + time_points1^2, factor.length = "adaptive.selection")
    
    
    
    var_bw <- NULL; var_remainder <- NULL
    mse_tr <-var_tr <- bias_tr <- NULL
    for(bw in seq(0.1,5,0.1)){
      tmp <- decompose_trend_remainder(x_trend_remainder, bw)
      var_bw <- c(var_bw, var(tmp$fitted_trend))
      mse_tr <- c(mse_tr, mean((tmp$fitted_trend -x_trend_remainder)^2))
      var_tr <- c(var_tr, var(tmp$fitted_trend -x_trend_remainder))
      bias_tr <- c(bias_tr, mean(tmp$fitted_trend -x_trend_remainder)^2)
      var_remainder <- c(var_remainder, mean( (tmp$fitted_remainder)^2))
    }
    
    
    #plot(seq(0.1,5,0.1),mse_tr)
    #plot(seq(0.1,5,0.1),bias_tr)
    
    
    par(mfrow = c(1,2))
    plot(seq(0.1,5,0.1),var_bw)
    plot(seq(0.1,5,0.1),var_remainder)
    
    #plot(var_bw+ var_remainder)
  }
  
}  
```











# DFT-IDFT to de-noise
```{r}
load(file = "D:/NSERC/project/1409 5-day avg.RData")

fivedaily = log(fivedaily_df$resp)

```


```{r}

s5 = seasonal(fivedaily, deg=2, window=38*12)
t5 = trend(fivedaily-s5, deg = 2, window = 0.1*length(fivedaily))
r5 = fivedaily - s5 - t5
acf(r5, lag.max = 500, na.action = na.exclude)
pacf(r5, na.action = na.exclude)

plot(fivedaily, col="green", ylim=c(-5,5))
lines(s5, col = "red")
lines(t5, col = "blue")


plot(fivedaily-s5, col = "green")
lines(t5, col = "blue")

```


```{r}
plot(r5)
model = arima(r5, order = c(3,0,0))
tsdiag(model)
auto.arima(r5)

```



```{r}
ft = fft(as.vector(fivedaily-s5)[1:1200], inverse = FALSE)

ft[9:(1200-9+2)] = 0+0i
plot((fivedaily-s5)[1:1200], col = "green")
lines(Re(fft(ft, inverse = TRUE)/1200), col = "blue")
```



```{r}
ft_s = fft(as.vector(s5)[1:1200], inverse = FALSE)
# ft_s[104:(1200-104+2)] = 0+0i

plot(as.vector(s5)[1:1200], type = "l", col = "blue")
lines(Re(fft(ft_s, inverse = TRUE)/1200), col= "red")
```



```{r}
plot(Re(ft)[1:600], type = "l", col = "blue")
plot(Re(ft_s)[1:600],type = "l", col = "blue")
# lines(abs(ft_s), col = "red")

```



```{r}
ft_sim = fft(as.vector(Xt), inverse = FALSE)
q = 5
ft_sim[q:(3650-q+2)] = 0+0i
plot(as.vector(Xt), col = "green", main = "res with ARMA(1,2), critfreq = 3")
lines(Re(fft(ft_sim, inverse = TRUE)/3650), type = "b", col = "blue")
lines(as.vector(tr), type = "b", col = "red")
lines(as.vector(trend(Xt, deg = 2, window = 0.3 * n)), type = "b", col = "black")
legend("bottomright",legend=c("data","true", "DFT", "loess"), lty=1,
       col=c("green","red", "blue", "black"))

```



```{r}
save.image(file = "simulation.RData")

```