---
title: "watershed uts utility"
author: "Junsong Tang"
date: '2022-07-24'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tseries)
library(xts)
library(zoo)
library(tidyverse)
library(forecast)
library(stlplus)
```




#=========================================================================================
# Format the time series
```{r, echo=TRUE}
# month-format
month.unify = function() {
  month = NULL
  for (m in (1:12)) {
    if ((m <= 7 && m %% 2 == 1) || (m >= 8 && m %% 2 == 0)) {
      month = c(month, rep(m, 31))
    } else if (m == 2) {
      month = c(month, rep(m, 28))
    } else {
      month = c(month, rep(m, 30))
    }
  }
  return (month)
}


# day-format
day.unify = function() {
  day = NULL
  for (m in (1:12)) {
    if ((m <= 7 && m %% 2 == 1) || (m >= 8 && m %% 2 == 0)) {
      day = c(day, seq(from = 1, to = 31, by = 1))
    } else if (m == 2) {
      day = c(day, seq(from = 1,to = 28,by = 1))
    } else {
      day = c(day, seq(from = 1,to = 30,by = 1))
    }
  }
  return (day)
}

```


# Daily structure
```{r, echo=TRUE}
# given a Rdata file, return a complete data frame(with covariates & response)
as.daily = function(data, impute=FALSE) {
  #start year
  start = data$yrs[1]
  # number of years
  n = length(data[[1]][, 1])
  
  # format unit month & days in calender
  month = month.unify()
  day = day.unify()
  # repeat unit format for n years
  months = rep(month, n)
  days = rep(day, n)
  # format years from 1980 to 
  years = rep(start:(start+n-1), each=365)
  
  frame = data.frame(years, months, days)
  
  # refined version
  for (k in (1:(length(data)-1))) {
    values = as.vector(t(data[[k]]))
    
    # # N/A & imputation if applicable
    # if (impute) {
    #   na_indices = which(is.na(values))
    #   # imputation method
    #   values[na_indices] = values[na_indices - 365]
    # }
    
    # some data file do not have covariate/resp lists of same length:
    # i.e. length(snow)=length(rain) != length(resp)
    if (dim(frame)[1] == length(values)) {
      frame = cbind(frame, values)
      names(frame)[k+3] = substr(names(data)[k], start=1, stop=4)
    }
    # # else, we could pad the shorter vector: length(values) = dim(frame)[1]
    # else {
    #   length(values) = dim(frame)[1]
    #   frame = cbind(frame, values)
    #   names(frame)[k+3] = names(data)[k]
    # }
  }
  
  # N/A & imputation if applicable
  if (impute) {
    l = dim(frame)[1]
    w = dim(frame)[2]
    for (i in (1:l)) {
      for (j in (4:w)) {
        if (is.na(frame[[j]][i])) {
          frame[[j]][i] = frame[[j]][i-365]
        }
      }
    }
  }
  
  
  
  # # original version
  # for (k in (1:(length(Rdata)-1))) {
  #   # data one of matrices in Rdata(cube) file
  #   mtx = Rdata[[k]]
  #   values = NULL
  #   for (i in (1:n)) {
  #     for (j in (1:length(mtx[1,]))) {
  #       values = c(values, mtx[i,j])
  #       
  #       #N/A & imputation if applicable
  #       if (impute) {
  #         ind = (i-1)*365 + j
  #         if (is.na(values[ind])) {
  #           values[ind] = values[ind-365]
  #         }
  #       }
  #       
  #     }
  #   }
  #   
  #   if (dim(frame)[1] == length(values)) {
  #     frame = cbind(frame, values)
  #     names(frame)[k+3] = names(Rdata)[k]
  #   }
  #   # else, we could pad the shorter vector: length(values) = dim(frame)[1]
  #   # frame = cbind(frame, values)
  #   # names(frame)[k+3] = names(Rdata)[k]
  # }
  return (frame)
}

```


# Monthly structure(smooth)
```{r, echo=TRUE}
as.monthly = function(data) {
  
  daily = as.daily(data, impute = FALSE)
  # start year
  start = daily$years[1]
  k = dim(daily)[2]-3 # total number of category(i.e. snow, rain, temp...)
  
  monthly = data.frame()
  
  year = NULL
  month = NULL
  
  m = 1 # month cursors
  sum = rep(0, k) # sum accumulator(row vector!)
  
  count = 0 # days accumulator
  ycount = 0 #  years accumulator
  for (i in (1:length(daily$months))) {
    # get all current values of snow, rain, temp, resp in vector
    curr = NULL
    for (j in (1:k)) {
      curr = c(curr, daily[[3+j]][i])
      # curr = c(curr, daily[i,(3+j)])
    }
    # curr = daily[i,][-(1:3)]: # this method is very slow!
    
    # sum: row vector
    # curr: row vector
    # avg: row vector
    # monthly: data frame/matrix
    
    # with the same month && not at December 31st
    if (m == daily$months[i] && i < length(daily$months)) {
      sum = sum + curr
      count = count + 1
    } # with the same month && at December 31st
    else if (m == daily$months[i] && i == length(daily$months)) {
      sum = sum + curr
      count = count + 1
      avg = sum / count
      monthly = rbind(monthly, avg)
      month = c(month, m)
      year = c(year, start+ycount)
    } # at the point of month change(1st day of next month)
    else {
      avg = sum / count
      monthly = rbind(monthly, avg)
      month = c(month, m)
      year = c(year, start+ycount)
      if (m == 12) {
        ycount = ycount + 1
      }
      m = daily$months[i]
      sum = curr
      count = 1
    }
  }
  # note k+3 = dim(daily)[2]
  names(monthly) = names(daily[4:(k+3)])
  return (data.frame(year, month, monthly))
}

```










#=======================================================================================
# Seasonal & Trend


(1) My method
# Seasonal
```{r, echo=TRUE}
# helper: re-arrange the raw time series into same day equivalent classes

# trim: boolean - whether to trim out the outliers of cluster(matrix row)

# same_day_cluster = function(ts, trim=FALSE) {
#   k = floor(tsp(ts)[2]-tsp(ts)[1])+1
#   freq = as.integer(tsp(ts)[3])
# 
#   M = list()
#   for (i in (1:freq)) {
#     indices = seq(from=i, to=i+(k-1)*freq, by=freq)
#     row = ts[indices]
#     # whether to remove outliers or not
#     if (trim) {
#       # set quantile & subset
#       range = quantile(row, prob = c(0.25, 0.75), na.rm = TRUE)
#       lb = as.numeric(range[1])
#       ub = as.numeric(range[2])
#       row = row[lb<=row&row<=ub]
#     }
#     M[[i]] = row
#   }
#   return (M)
# }


# trim: percent of lower tail
same_day_cluster = function(ts, trim=0) {
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1
  freq = as.integer(tsp(ts)[3])
  
  M = list()
  for (i in (1:freq)) {
    indices = seq(from=i, to=i+(k-1)*freq, by=freq)
    row = ts[indices]
    # whether to remove outliers or not
    if (trim != 0) {
      # set quantile & subset
      range = quantile(row, prob = c(trim, 1-trim), na.rm = TRUE)
      lb = as.numeric(range[1])
      ub = as.numeric(range[2])
      row = row[lb<=row&row<=ub]
    }
    M[[i]] = row
  }
  return (M)
}



# trim: real number in [0,1], representing percentage of the lower tail
my.seasonal = function(ts, window, degree, trim) {
  # number of years
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1
  freq = as.integer(tsp(ts)[3])
  start_year = as.integer(tsp(ts)[1])
  
  equi_class = same_day_cluster(ts, trim=trim)
  n = length(equi_class[[1]])
  
  seasonality = NULL
  for (i in (1:freq)) {
    # [-w+i,..i,...w+i]
    interval = seq(from=-window+i, to=window+i, by=1)
    y = NULL
    x = NULL
    for (w in interval) {
      if (w < 1) {
        cluster = (equi_class[[w+freq]])[-n]
      }
      else if (w > freq) {
        cluster = (equi_class[[w-freq]])[-1]
      }
      else {
        cluster = equi_class[[w]]
      }
      
      cluster = cluster[!is.na(cluster)]
      
      y = c(y, cluster)
      x = c(x, rep(w-i, length(cluster)))
    }
    
    if (degree == 0) {
      model = lm(y ~ 1)
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 1) {
      model = lm(y ~ x)
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 2) {
      model = lm(y ~ x+I(x^2))
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    if (degree == 3) {
      model = lm(y ~ x+I(x^2)+I(x^3))
      yhat = as.numeric(model$fitted.values[which(x==0)][1])
    }
    
    slots = seq(from=i, to=i+(k-1)*freq, by=freq)
    seasonality[slots] = yhat
  }
  
  return (ts(seasonality, start=c(start_year, 1), frequency = freq))
}

```


# Cross validation(5 fold by default)
# 38 => 8 + 8 + 8 + 8 + 6
```{r, echo=TRUE}
cv.myseason = function(ts, window.max = 10, trim_per.max = 0.25, fold = 5) {
  start_year = tsp(ts)[1]
  end_year = tsp(ts)[2]
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1 #k = 38 years
  t = ceiling(k/fold) #t = 8 years/unit
  freq = as.integer(tsp(ts)[3])
  minSMSR = Inf
  best_window = 0
  best_degree = 0
  best_trim = 0
  for (w in (0:window.max)) {
    for (d in (0:3)) {
      for (p in (0:(trim_per.max/0.05))) {
        # sum of mean squared error over CVs
        SMSR = 0
        # run fold times CV
        for (i in (1:fold)) {
          test_index = seq(from=1+(i-1)*t*freq, to=min(i*t*freq, k*freq), by=1)
          test = ts[test_index]
          test = ts(test, start=c(start_year+(i-1)*t, 1), frequency = freq)
          
          train = ts[-test_index]
          train = ts(train, start=c(start_year, 1), frequency = freq)
          seasonality = my.seasonal(train, w, d, trim=p*0.05)
          
          
          pred = seasonality[1:length(test)]
          residual = (test - pred)
          SMSR = SMSR + mean(residual^2, na.rm = TRUE)
        }
        
        
        
        # for (year in (1:k)) {
        #   test_index = seq(from=1+(year-1)*freq, to=year*freq, by=1)
        #   test = ts[test_index]
        #   # test must be converted into ts object!!!
        #   test = ts(test, start=c(start_year+(year-1), 1), frequency = freq)
        #   train = ts[-test_index]
        #   # train must be converted into ts object!!!
        #   train = ts(train, start=c(start_year, 1), frequency = freq)
        #   seasonality = seasonality4(train, w, d)
        #   pred = seasonality[1:freq]
        #   residual = (test - pred)
        #   SMSR = SMSR + mean(residual^2, na.rm = TRUE)
        # }
        
        if (SMSR < minSMSR) {
          best_window = w
          best_degree = d
          best_trim = p
          minSMSR = SMSR
        }
      }
    }
  }
  
  return (data.frame(best_window, best_degree, best_trim))
}

```

# trend(MA over one full period)
```{r, echo=TRUE}
MA.smooth = function(x, lag) {
  n = length(x)
  result = NULL
  for (i in (1:(n-lag))) {
    avg = mean(x[i:(i+lag)], na.rm = TRUE)
    result = c(result, avg)
  }
  return (result)
}


my.trend = function(ts) {
  freq = as.integer(tsp(ts)[3])
  n = length(ts)
  start_year = as.integer(tsp(ts)[1])
  
  trend = NULL
  final_trend = rep(NA, n)
  # freq=5, w=2, freq=4, w=1.5
  window  = (freq-1) / 2
  # if freq = odd
  if (freq %% 2 != 0) {
    trend = MA.smooth(ts, lag=2*window)
    # padding head & tail
    final_trend[(window+1):(n-window)] = trend
  }
  else {
    temp = MA.smooth(ts, lag=2*window)
    trend = MA.smooth(temp, lag = 1)
    # padding head & tail
    w = window + 0.5
    final_trend[(w+1):(n-w)] = trend
  }
  
  return (ts(final_trend, start=c(start_year, 1), frequency = freq))
  
}
```











(2) LOESS method
# Seasonal
```{r, echo=TRUE}
seasonal.loess = function(ts, deg=2, span=0.75, iter=4) {
  k = floor(tsp(ts)[2]-tsp(ts)[1])+1
  freq = as.integer(tsp(ts)[3])
  start_year = as.integer(tsp(ts)[1])
  # form data frame arranged in same day order
  x = NULL
  y = NULL
  z = NULL
  for (i in (1:freq)) {
    indices = seq(from=i, to=i+(k-1)*freq, by=freq)
    cluster = ts[indices]
    # clear out NA
    cluster = cluster[!is.na(cluster)]
    x = c(x, rep(i, length(cluster)))
    y = c(y, cluster)
    # set up marker such that it represents the pos that we should extract the fitted value 
    z = c(z, TRUE, rep(FALSE, length(cluster)-1))
    
  }
  # df = data.frame(x, y)
  
  loess_obj = loess(y ~ x, span=span, degree=deg, family="symmetric", method="loess", control =loess.control(iterations=iter))
  
  full_period = loess_obj$fitted[which(z==TRUE)]
  seasonality = rep(full_period, k)
  
  return (ts(seasonality, start=c(start_year,1), frequency = freq))
}

```


# Trend
```{r, echo=TRUE}
trend.loess = function(ts, deg=2, span=0.75, iter=4, impute=FALSE) {
  freq = as.integer(tsp(ts)[3])
  start_year = as.integer(tsp(ts)[1])
  n = length(ts)
  y = ts
  x = c(1:n)
  loess_obj = loess(y ~ x, na.action=na.exclude, span=span, degree=deg, family="symmetric", method="loess", control=loess.control(surface="direct", iterations=iter))
  na_index = which(is.na(ts))
  y[-na_index] = loess_obj$fitted
  if (impute) {
    y[na_index] = as.numeric(predict(object=loess_obj, newdata=data.frame(x=na_index)))
  }
  else {
    y[na_index] = NA
  }
  
  return (ts(y, start = c(start_year,1), frequency = freq))
  
}

```




(3) Convergence test(alternate season & trend extraction difference convergence test)
$$X - s_1 = t_1 + \varepsilon_1$$
$$X - t_1 = s_2 + \varepsilon_2$$
$$X - s_2 = t_2 + \varepsilon_3$$
$$\ldots$$
$$X - s_{n-1} = t_{n-1} + \varepsilon_{2n-3}$$
$$X - t_{n-1} = s_n + \varepsilon_{2n-2}$$
$$X - S_n = t_n + \varepsilon_{2n-1}$$

```{r, echo=TRUE}
#iter should be odd
convergence.test = function(ts, iter) {
  ds = NULL
  dt = NULL
  prev_season = rep(0, length(ts))
  prev_trend = rep(0, length(ts))
  for (i in (1:iter)) {
    if (i == 1) {
      curr_season = seasonal.loess(ts, span=1/30)
      curr_trend = trend.loess(ts - curr_season, span=1/38)
      ds = c(ds, mean((curr_season - prev_season)^2, na.rm=TRUE))
      dt = c(dt, mean((curr_trend - prev_trend)^2, na.rm=TRUE))
      prev_season = curr_season
      prev_trend = curr_trend
    }
    else if (i %% 2 != 0) {
      curr_trend = trend.loess(ts - prev_season, span=1/38)
      dt = c(dt, mean((curr_trend - prev_trend)^2, na.rm=TRUE))
      prev_trend = curr_trend
    }
    else {
      curr_season = seasonal.loess(ts - prev_trend, span=1/30)
      ds = c(ds, mean((curr_season - prev_season)^2, na.rm=TRUE))
      prev_season = curr_season
    }
  }
  
  return(data.frame(ds, dt))
}

```



#relative error = sum(r_old^2) - sum(r_new^2)/ sum(r_new^2)

#do while(sum(r_old^2) - sum(r_new^2) < 1 - relative error)

$$\displaystyle r = \frac{|\sum_{i=1}^m \varepsilon_{n,i} - \sum_{i=1}^m \varepsilon_{n-1,i}|}{\sum_{i=1}^m \varepsilon_{n,i}}$$

```{r, echo=TRUE}
residual.test = function(ts, iter) {
  relative_errors = NULL

  prev_season = rep(0, length(ts))
  prev_trend = rep(0, length(ts))
  prev_error = rep(0, length(ts))
  for (i in (1:iter)) {
    if (i == 1) {
      curr_season = seasonal.loess(ts, span=1/30)
      curr_trend = trend.loess(ts - curr_season, span=1/38)
      curr_error = ts - curr_season - curr_trend
      
      curr = sum(curr_error^2, na.rm = TRUE)
      prev = sum(prev_error^2, na.rm = TRUE)
      
      relative_errors = c(relative_errors, abs(curr - prev)/curr)
      
      prev_season = curr_season
      prev_trend = curr_trend
      prev_error = curr_error
      
    }
    else if (i %% 2 != 0) {
      curr_trend = trend.loess(ts - prev_season, span=1/38)
      curr_error = ts - curr_trend - prev_season
      curr = sum(curr_error^2, na.rm = TRUE)
      prev = sum(prev_error^2, na.rm = TRUE)
      
      relative_errors = c(relative_errors, abs(curr - prev)/curr)

      prev_trend = curr_trend
      prev_error = curr_error
    }
    else {
      curr_season = seasonal.loess(ts - prev_trend, span=1/30)
      curr_error = ts - curr_season - prev_trend
      curr = sum(curr_error^2, na.rm = TRUE)
      prev = sum(prev_error^2, na.rm = TRUE)
      
      relative_errors = c(relative_errors, abs(curr - prev)/curr)

      prev_season = curr_season
      prev_error = curr_error
    }
  }
  
  return(relative_errors)
}

```





(4) Cross Validation
$$\text{total mean squrared error} = \frac1n \sum_{i=1}^{n} \varepsilon_{i}^2$$
$$\text{In k-fold CV, }\sum_{j=1}^k n_j = n$$

$$\text{span.max} = \frac{k}{\sqrt{n}}$$

# common suggestion of truncation point M = 2*sqrt(N);(M<N)
#cv.seasonal
```{r, echo=TRUE}
cv.seasonal = function(ts, d.max=2, span.min=NULL, span.max=NULL, span.control=NULL, fold=5) {
  start_year = tsp(ts)[1]
  end_year = tsp(ts)[2]
  k = floor(end_year-start_year)+1 #k = 38 years
  t = ceiling(k/fold) #t = t years/unit
  freq = as.integer(tsp(ts)[3])
  minTMSE = Inf

  best_degree = 0
  best_span = span.min
  
  n = length(ts)
  if (is.null(span.min)) {
    span.min = 10/sqrt(n)
  }
  
  if (is.null(span.max)) {
    span.max = 20/sqrt(n)
  }
  
  if (is.null(span.control)) {
    span.control = (span.max - span.min) / 10
  }
  
  for (d in (1:d.max)) {
    for (s in seq(from=span.min, to=span.max, by=span.control)) {
      whole_pred = NULL
      # run 5 times CV
      for (i in (1:fold)) {
        test_index = seq(from=1+(i-1)*t*freq, to=min(i*t*freq, k*freq), by=1)
        test = ts[test_index]
        test = ts(test, start=c(start_year+(i-1)*t, 1), frequency = freq)
        
        train = ts[-test_index]
        train = ts(train, start=c(start_year, 1), frequency = freq)
        seasonality = seasonal.loess(train, deg=d, span=s)
        
        
        pred = seasonality[1:length(test)]
        whole_pred[test_index] = pred

      }
      TMSE = mean((ts - whole_pred)^2, na.rm=TRUE)
      
      
      if (TMSE < minTMSE) {
        best_degree = d
        best_span = s
        minTMSE = TMSE
      }
    }
  }
  
  return (data.frame(best_degree, best_span, minTMSE))
}



```





#cv.trend
```{r, echo=TRUE}
cv.trend = function(ts, d.max = 2, span.min=NULL, span.max=NULL, span.control=NULL, fold=5) {
  start_year = tsp(ts)[1]
  end_year = tsp(ts)[2]
  k = floor(end_year-start_year)+1 #k = 38 years
  t = ceiling(k/fold) #t = t years/unit
  freq = as.integer(tsp(ts)[3])
  n = length(ts)
  
  # total mean squared error 
  minTMSE = Inf

  best_degree = 0
  best_span = span.min
  
  n = length(ts)
  if (is.null(span.min)) {
    span.min = 0.1/sqrt(n)
  }
  
  if (is.null(span.max)) {
    span.max = 1/sqrt(n)
  }
  
  if (is.null(span.control)) {
    span.control = (span.max - span.min) / 10
  }
  
  for (d in (1:d.max)) {
    for (s in seq(from=span.min, to=span.max, by=span.control)) {
      whole_pred = NULL
      # run 5 times CV
      remain_index = c(1:n)
      for (i in (1:fold)) {
        # random_index = sample(c(1:length(remain_index)), size=min(ceiling(n/fold), length(remain_index)), replace=FALSE)
        # test_index = remain_index[random_index]
        
        test_index = sample(remain_index, size=min(ceiling(n/fold), length(remain_index)), replace=FALSE)
        
        
        # train_index = c(1:n)[-test_index]
        
        # test = ts[test_index]
        
        train = ts
        train[test_index] = NA
        train = ts(train, start=c(start_year,1), frequency=freq)
        
        trend = trend.loess(train, deg=d, span=s, impute=TRUE)
        pred = trend[test_index]
        whole_pred[test_index] = pred
        
        # remain_index = remain_index[-random_index]
        remain_index = setdiff(remain_index, test_index)
      }
      
      TMSE = mean((ts-whole_pred)^2, na.rm=TRUE)
      
      if (TMSE < minTMSE) {
        best_degree = d
        best_span = s
        minTMSE = TMSE
      }
    }
  }
  
  return (data.frame(best_degree, best_span, minTMSE))
}


```








#=====================================================================================
# p = TSA::periodogram(window(ts1409, start=c(1980, 1), end=c(1996,12)), xlim = c(0,0.01))
# d = data.frame(freq=p$freq, spec=p$spec)
# order = d[order(-d$spec),]
# top5 = head(order, 5)
# 1/top5$freq

```{r, echo=TRUE}
s = spectrum(ts1409, na.action=na.exclude, xlim = c(0,0.01), ylim=c(10, 1000), type = "h")
d = data.frame(freq=s$freq, spec=s$spec)
order = d[order(-d$spec),]
top5 = head(order, 5)
```

