---
title: "watershed"
author: "Junsong Tang"
date: "6/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(terra)
library(leaflet)
library(knitr)
library(tseries)
library(xts)
library(zoo)
library(forecast)
library(sp)
library(sf)
library(rgdal)
library(raster)
library(maptiles)
library(shiny)
library(ggplot2)
```



#========================================================================================# Global variables
```{r, echo=TRUE}
PATH = "D:/NSERC/project/processedData/"
FILENAMES = list.files(PATH)

# load shapes
SHAPES = vect("D:/NSERC/project/WaterShedDataExample/ToyData/Shape_Files/CANUS_Watershed_Boundaries.shp")


# shape_attr = st_read("D:/NSERC/project/WaterShedDataExample/ToyData/Shape_Files/CANUS_Watershed_Boundaries.shp")

SHAPE_attr = read_sf("D:/NSERC/project/WaterShedDataExample/ToyData/Shape_Files/CANUS_Watershed_Boundaries.shp")

# coordinates
GRIDINFO = read.csv("D:/NSERC/project/WaterShedDataExample/GridInfo.csv", sep = ";")


# given file path, return list of Rdata(cube)
loadAllData = function(path) {
  fileNames = list.files(path)
  filesList = list()
  for (i in (1:length(fileNames))) {
    suffix = fileNames[i]
    rdata = mget(load(paste0(path, suffix)))
    # mget returns a list of single-element list:[[*], [*], ...], use index form datafile[[i]]
    filesList[[i]] = rdata
  }
  return(filesList)
}



# # alldata is a list of Rdata(cube)
# alldata = loadAllData("D:/NSERC/project/processedData/")

```


# accessing shape file/Rdata file/grid code
```{r, echo=TRUE}
GRIDCODES = GRIDINFO$gridcode

shapeIDs = SHAPE_attr$gridcode
# length(GRIDCODES) == length(shapeIDs)
# length(GRIDCODES) == length(GRIDCODES == shapeIDs)
# conclusion: gridInfo has the same order and length of shapes, in fact we can use gridInfo index to access single shape file in shapes


# 1409 Rdata file and its corresponding shape file
shape1409 = SHAPES[which(GRIDCODES == 1409)]
# 1409 Rdata file(list of matrices)
data1409 = mget(load(paste0(PATH, "1409_dat.Rdata"))) 
```





#========================================================================================
# Helpers(under construction)

# ts format
```{r, echo=TRUE}
# univariate time series
# data: matrix or vector/list form containing all daily data of single category(snow,rain)
# returns ts object from start_year to end_year with freq
uts = function(data, freq = 365, start_year, end_year) {
  if (is.vector(data) || is.list(data)) {
    return (ts(data, start = c(start_year, 1), end = c(end_year, 365), frequency = freq))
  }
  if (is.matrix(data)) {
    # as.vector goes along the column(vertically) of a matrix
    y = as.vector(t(data))
    return (ts(y, start = c(start_year, 1), end = c(end_year, 365), frequency = freq))
  }
  
  
}

# # category is string: the matrix name in Rdata file, i.e. "resp_mat", "rain_mat"
# # returns ts object of given category
# cube_to_ts = function(Rdata, category) {
#   mtx = Rdata[[which(names(Rdata) == category)]]
#   start = Rdata$yrs[1]
#   end = Rdata$yrs[length(Rdata$yrs)]
#   return (matrix_to_ts(mtx, start_year=start, end_year=end))
# }


# multivariate time series
# input: list form of watershed data, i.e. Rdata, data frame
# return a multivariate ts in data frame
mts = function(data) {
  df = data.frame()
  for (i in (1:(length(data)-1))) {
    mtx = data[[i]]
    start = data$yrs[1]
    end = data$yrs[length(data$yrs)]
    ts = uts(mtx, start_year=start, end_year=end)
    if (dim(df)[1] == 0) {
      df = data.frame(ts)
    } # if current ts has same length as previous ones
    else if (dim(df)[1] == length(ts)) {
      df = cbind(df, ts)
    } # if current ts differs in length from previous ones, then trim(longer)/pad(shorter)
    else {
      length(ts) = dim(df)[1]
      df = cbind(df, ts)
    }
    names(df)[i] = substr(names(data)[i],start=1,stop=4)
  }
  return (df)
  
}

# remark: mts is data frame, also list object, to access one ts element, use list index


```

#test
```{r, echo=TRUE}
start_time = Sys.time()
mts1409 = mts(data1409)
end_time = Sys.time()
end_time - start_time

```






Operations by Watershed object
=========================================================================================
# Watershed object

attributes:
# mts(snow, rain, temp, resp)
# gridcode
# index: shape object index
# name: name of the watershed


method:
# get_fraction()
# render_plot
# ...


constructor input:
#data: has to be in list form(data frame containing covariates), because it has to contain snow, rain, temp, resp
# (i.e. RData, data frame), also could be null
# gridcode has to be integer, non-null
# index has to be integer, non-null
```{r, echo=TRUE}
Watershed = function(data, gridcode, index) {
  # check shape & gridcode
  # (1) all inputs are correct
  shape = SHAPES[index]
  if (shape$gridcode == gridcode && is.numeric(gridcode) && gridcode %in% GRIDCODES){
    obj = NULL
    if (is.null(data)) {
      mts = NULL
    }
    else if ((is.vector(data) || is.list(data) || is.data.frame(data))) {
      mts = mts(data)
    }
    else {
      stop("data is not of correct type!")
      return ()
    }
    obj$mts = mts
    obj$gridcode = gridcode
    obj$index = index
    obj$name = shape$Station_NA
    class(obj) = "Watershed"
    return (obj)
    
  }
  else if (shape$gridcode != gridcode){
    stop("gridcode mismatch shape!")
    return ()
  } 
  else if (!is.numeric(gridcode)) {
    stop("gridcode is not of correct type!")
    return ()
  }
  else {
    stop("such grid code does not exist!")
    return ()
  }
  
}


```


```{r, echo=TRUE}
start_time = Sys.time()
ws1409 = Watershed(data1409, 1409, 1368)
end_time = Sys.time()
end_time-start_time

```











=========================================================================================
# TODO (under construction)
# Arithmetic
```{r, echo=TRUE}
# snow fraction
# input: Watershed object
# return: snow_fraction(double)
snow_fraction = function(ws) {
  if (is.null(ws$mts)) {
    return (NA)
  }
  else {
    snow = ws$mts$snow
    rain = ws$mts$rain
    snow_sum = 0
    rain_sum = 0
    for (i in (1:length(snow))) {
      s = snow[i]
      r = rain[i]
      # both snow and rain have to be non NA!
      if (!is.na(s) && !is.na(r)) {
        snow_sum = snow_sum + s
        rain_sum = rain_sum + r
      }
    }
    snow_ratio = snow_sum / (snow_sum + rain_sum)
    return (snow_ratio)
  }
}

# rain fraction
# input: Watershed object
rain_fraction = function(ws) {
  return (1 - snow_fraction(ws))
}

```


```{r, echo=TRUE}
start_time = Sys.time()
snow_fraction(ws1409)
end_time = Sys.time()
end_time-start_time


```




#========================================================================================
# Functionalities

# TODO (under construction)
# fetch summary
```{r, echo=TRUE}
# input: watersheds = list(ws1, ws2...) ws watershed object
get_fraction = function(watersheds) {
  snow_fraction = NULL
  rain_fraction = NULL
  gridcode = NULL
  for (ws in watersheds) {
    gridcode = c(gridcode, ws$gridcode)
    snow_fraction = c(snow_fraction, snow_fraction(ws))
    rain_fraction = c(rain_fraction, rain_fraction(ws))
  }
  return (data.frame(gridcode, snow_fraction, rain_fraction))
  
}


```

```{r, echo=TRUE}
start_time = Sys.time()
get_fraction(watersheds = list(ws1409))
end_time = Sys.time()
end_time-start_time


```

# Archive

<!-- ```{r, echo=TRUE} -->
<!-- get_fraction = function(shapes) { -->
<!--   grid_codes = shapes$gridcode -->
<!--   snow_ratios = NULL -->
<!--   rain_ratios = NULL -->
<!--   for (i in (1:length(grid_codes))) { -->
<!--     filename = paste0(grid_codes[i], "_dat.RData") -->
<!--     # exception handling -->
<!--     if (!(filename %in% fileNames)) { -->
<!--       snow_ratios = c(snow_ratios, NA) -->
<!--       rain_ratios = c(rain_ratios, NA) -->
<!--     } -->
<!--     else { -->
<!--       data = mget(load(paste0(path, grid_codes[i], "_dat.RData"))) -->
<!--       sr = snow_fraction(data) -->
<!--       rr = rain_fraction(data) -->
<!--       snow_ratios = c(snow_ratios, sr) -->
<!--       rain_ratios = c(rain_ratios, rr) -->
<!--     } -->
<!--   } -->
<!--   return (data.frame(grid_codes, snow_ratios, rain_ratios)) -->
<!-- } -->
<!-- ``` -->


<!-- #test -->
<!-- ```{r, echo=TRUE} -->
<!-- # indices = c(1061, 1273, 1368) -> watershed file:(1003, 1013, 1409) -->
<!-- get_fraction(shapes[c(1061,1273,1368)]) -->
<!-- render_piechart(shapes[c(1061,1273,1368)]) -->

<!-- ``` -->




```{r, echo=TRUE}
# TODO
# gridcodes: vector of gridcode, i.e.c(1003, 1013, 1409), c(1,10,1001,1409)
# find_watersheds = function() {}
# search by indices: another option
load_watersheds = function(gridcodes) {
  watersheds = list()
  for (i in (1:length(gridcodes))) {
    gc = gridcodes[i]
    filename = paste0(gc, "_dat.RData")
    if (!(filename %in% FILENAMES)) {
      data = NULL
    } else {
      data = mget(load(paste0(PATH, filename)))
    }
    index = which(SHAPES$gridcode == gc)
    ws = Watershed(data, gc, index)
    watersheds[[i]] = ws
  }
  return (watersheds)
  
}



```





# TODO (under construction)
# load any info about the watersheds/shapes
```{r, echo = TRUE}
# input: list of Watershed object
# return: content => list of HTML string
show_fraction = function(watersheds) {
  content = NULL
  for (i in (1:length(watersheds))) {
    stats = get_fraction(watersheds)
    sr = stats[i,][2]
    rr = stats[i,][3]
    if (is.na(sr) || is.na(rr)) {
      content = c(content, paste0("<br>watershed:", stats[i,][1], 
                                  "<br>snow fraction:", stats[i,][2],
                                  "<br>rain fraction:", stats[i,][3]))
    } else {
      content = c(content, paste0("<br>watershed:", stats[i,][1], 
                                  "<br>snow fraction:", round(stats[i,][2],2)*100,"%",
                                  "<br>rain fraction:", round(stats[i,][3],2)*100,"%"))
    }
  }
  return (content)
}

# input: list of Watershed object
# return: content => list of HTML string
show_ts = function(watersheds) {
  content = NULL
  for (i in (1:length(watersheds))) {
    if (!is.null(watersheds[[i]])) {
      
    } else {
      content = c(content, "<br>no data")
    }
    
  }
  return (content)
}

# input: list of Watershed object
# return: content => list of HTML plot
show_plot = function(watersheds, option) {
  plot = NULL
  
  return (plot)
}

load_content = function(watersheds, option) {
  content = NULL
  switch(option,
         "fraction" = {content = show_fraction(watersheds)},
         "time series" = {content = show_ts(watersheds)},
         "piechart" = {content = show_plot(watersheds)}
         
  )
  return (content)
  
}

```


# TODO (under construction)
# plot watersheds's boundaries(shapes)
```{r, echo=TRUE}
# shapes: a set of shape objects, e.g., shp_dat[1:3];
# content: 
# map: boolean - TRUE puts boundaries on a map;
plot_shape = function(shapes, content=NULL, map=TRUE, option="fraction") {
  # center_longs = shapes$Longitude
  # center_lats = shapes$Latitude
  # list of x coord(longitude) of centroid
  center_longs = NULL
  # list of y coord(latitude) of centroid
  center_lats = NULL
  for (i in (1:length(shapes))) {
    center_longs = c(center_longs, mean(geom(shapes[i])[,3]))
    center_lats = c(center_lats, mean(geom(shapes[i])[,4]))
  }
  
  if (map) {
    # leaflet(sf::st_as_sf(shapes)) %>% addTiles() %>% addPolygons(fillOpacity=0.1) %>% addMarkers(lng=center_longs, lat=center_lats, popup=content) 
    # %>% addPopups(center_longs, center_lats, popup=content, options = popupOptions(closeButton = TRUE))
    m = leaflet(sf::st_as_sf(shapes))
    m = addTiles(m) 
    m = addPolygons(m, fillOpacity=0.1)
    m = addMarkers(m, lng=center_longs, lat=center_lats, popup=content)
    # # mouse hovering to get content
    # m = addMarkers(m, lng=center_longs, lat=center_lats, label=lapply(content, HTML))
    m
  }
  else {
    plot(shapes)
    # m = leaflet(sf::st_as_sf(shapes))
    # m = addPolygons(m, fillOpacity=0.1)
    # m = addMarkers(m, lng=center_longs, lat=center_lats, popup=content)
    # m
  }
}


# watersheds: list of watershed objects
plot_watersheds = function(watersheds, map = TRUE, option = "fraction") {
  content = load_content(watersheds, option)
  indices = sapply(watersheds, FUN=function(ws){ws$index})
  shapes = SHAPES[indices]
  plot_shape(shapes, content, map, option)
}

```

#tests
```{r, echo=TRUE}
# plot_shape(SHAPES[1:4], FALSE, "fraction")
# c1 = load_content(ws_1_2_1409,option = "fraction")
# plot_shape(SHAPES[1:4], map=TRUE, "fraction")
# plot_shape(SHAPES[c(1, 11, 101, 31, 77)], TRUE, "fraction")
# plot_shape(SHAPES[c(1061,1273,1368)], map=TRUE, "fraction")
# plot_shape(SHAPES[c(1061,1273,1368)], map=FALSE, "fraction")

plot_watersheds(watersheds=load_watersheds(c(1003,1013,1409)), map=TRUE)
plot_watersheds(watersheds=load_watersheds(c(1, 77, 101, 31)), map=TRUE)

```


























# inside region?
```{r, echo=TRUE}
# shape: a single shape file
# long_min, long_max, lat_min, lat_max: lower and upper bounds on longitude and latitude;
# Returns a boolean whether the shape is wholly inside the region or just partially inside;;
inside_region = function(shape, long_min, long_max, lat_min, lat_max) {
  x_min = xmin(shape)
  x_max = xmax(shape)
  y_min = ymin(shape)
  y_max = ymax(shape)
  return (long_min <= x_min && x_min <= long_max && lat_min <= y_min && y_max <= lat_max)
  # longs = geom(shape)[,3]
  # lats = geom(shape)[,4]
  # strict = TRUE
  # i = 1
  # while (strict && i <= length(longs)) {
  #   if (!(long_min <= longs[i] && longs[i] <= long_max && lat_min <= lats[i] && lats[i] <= lat_max)) {
  #     strict = FALSE
  #   }
  #   i = i + 1
  # }
  # return (strict)
}


```


#tests
```{r, echo=TRUE}
plot_shape(SHAPES[1], map=FALSE)
start_time = Sys.time()
inside_region(SHAPES[1], -132, -125, 57, 59.5)
inside_region(SHAPES[1], -133, -126, 57, 59.5)
end_time = Sys.time()
end_time - start_time

```





# in_which_region
```{r, echo=TRUE}
# shapes: a list of shape files
# Returns the index of the shape containing the longitude and latitude
in_which_region = function(shapes, long, lat) {
  indices = NULL
  for (i in (1:length(shapes))) {
    longs = geom(shapes[i])[,3]
    lats = geom(shapes[i])[,4]
    if (point.in.polygon(point.x = long, 
                         point.y = lat,
                         pol.x = longs,
                         pol.y = lats)) {
      indices = c(indices, i)
    }
  }
  return (indices)
}

```


#tests
```{r, echo=TRUE}
plot_shape(SHAPES[1:4], FALSE)
in_which_region(SHAPES[1:4], long = -125, lat = 55.1)
```


# in_which_region(return grid code/station ID)
```{r, echo=TRUE}
# in_which_region = function(shapes, long, lat) {
#   IDs = NULL
#   for (i in (1:length(shapes))) {
#     longs = geom(shapes[i])[,3]
#     lats = geom(shapes[i])[,4]
#     if (point.in.polygon(point.x = long, 
#                          point.y = lat,
#                          pol.x = longs,
#                          pol.y = lats)) {
#       IDs = c(IDs, shapes[i]$Station_ID)
#     }
#   }
#   return (IDs)
# }

```





```{r, echo=TRUE}
# shapes[1:2] is "SpatVector" object
# ext(shapes[1:2]) is "SpatExtent" object
# box is "SpatRaster" object
# Http error 404 can be solved by giving different online map provider
box = get_tiles(x = ext(SHAPES[1:3]), provider = "Esri.WorldStreetMap")
plotRGB(box)
lines(SHAPES[1:3], col = "blue")
lines(ext(SHAPES[1:3]))


```









# Format(temporary finished refinement)
```{r, echo=TRUE}
# month-format
month_format = function() {
  month = NULL
  for (m in (1:12)) {
    if ((m <= 7 && m %% 2 == 1) || (m >= 8 && m %% 2 == 0)) {
      month = c(month, rep(m, 31))
    } else if (m == 2) {
      month = c(month, rep(m, 28))
    } else {
      month = c(month, rep(m, 30))
    }
  }
  return (month)
}


# day-format
day_format = function() {
  day = NULL
  for (m in (1:12)) {
    if ((m <= 7 && m %% 2 == 1) || (m >= 8 && m %% 2 == 0)) {
      day = c(day, seq(from = 1, to = 31, by = 1))
    } else if (m == 2) {
      day = c(day, seq(from = 1,to = 28,by = 1))
    } else {
      day = c(day, seq(from = 1,to = 30,by = 1))
    }
  }
  return (day)
}

# given a Rdata file, return a complete data frame(with covariates & response)
daily_frame = function(Rdata, impute=FALSE) {
  #start year
  start = Rdata$yrs[1]
  # number of years
  n = length(Rdata[[1]][, 1])
  
  # format unit month & days in calender
  month = month_format()
  day = day_format()
  # repeat unit format for n years
  months = rep(month, n)
  days = rep(day, n)
  # format years from 1980 to 
  years = rep(start:(start+n-1), each=365)
  
  frame = data.frame(years, months, days)
  
  # refined version
  for (k in (1:(length(Rdata)-1))) {
    values = as.vector(t(Rdata[[k]]))
    
    # # N/A & imputation if applicable
    # if (impute) {
    #   na_indices = which(is.na(values))
    #   # imputation method
    #   values[na_indices] = values[na_indices - 365]
    # }
    
    # some Rdata file do not have covariate/resp lists of same length:
    # i.e. length(snow)=length(rain) != length(resp)
    if (dim(frame)[1] == length(values)) {
      frame = cbind(frame, values)
      names(frame)[k+3] = substr(names(Rdata)[k], start=1, stop=4)
    }
    # # else, we could pad the shorter vector: length(values) = dim(frame)[1]
    # else {
    #   length(values) = dim(frame)[1]
    #   frame = cbind(frame, values)
    #   names(frame)[k+3] = names(Rdata)[k]
    # }
  }
  
  # N/A & imputation if applicable
  if (impute) {
    l = dim(frame)[1]
    w = dim(frame)[2]
    for (i in (1:l)) {
      for (j in (4:w)) {
        if (is.na(frame[[j]][i])) {
          frame[[j]][i] = frame[[j]][i-365]
        }
      }
    }
  }
  
  
  
  # # original version
  # for (k in (1:(length(Rdata)-1))) {
  #   # data one of matrices in Rdata(cube) file
  #   mtx = Rdata[[k]]
  #   values = NULL
  #   for (i in (1:n)) {
  #     for (j in (1:length(mtx[1,]))) {
  #       values = c(values, mtx[i,j])
  #       
  #       #N/A & imputation if applicable
  #       if (impute) {
  #         ind = (i-1)*365 + j
  #         if (is.na(values[ind])) {
  #           values[ind] = values[ind-365]
  #         }
  #       }
  #       
  #     }
  #   }
  #   
  #   if (dim(frame)[1] == length(values)) {
  #     frame = cbind(frame, values)
  #     names(frame)[k+3] = names(Rdata)[k]
  #   }
  #   # else, we could pad the shorter vector: length(values) = dim(frame)[1]
  #   # frame = cbind(frame, values)
  #   # names(frame)[k+3] = names(Rdata)[k]
  # }
  return (frame)
}



```


<!-- ```{r, echo=TRUE} -->
<!-- start_time = Sys.time() -->
<!-- df = daily_frame(data1409, impute = TRUE) -->
<!-- df[is.na(df$resp_mat),] -->
<!-- end_time = Sys.time() -->
<!-- end_time - start_time -->
<!-- ``` -->

<!-- ```{r, echo=TRUE} -->
<!-- start_time = Sys.time() -->
<!-- df = daily_frame(data1409, impute = FALSE) -->
<!-- df[is.na(df$resp_mat),] -->
<!-- end_time = Sys.time() -->
<!-- end_time - start_time -->
<!-- ``` -->



# monthly smooth
```{r, echo=TRUE}

monthly_average = function(Rdata) {
  
  daily = daily_frame(Rdata, impute = FALSE)
  # start year
  start = daily$years[1]
  k = dim(daily)[2]-3 # total number of category(i.e. snow, rain, temp...)
  
  monthly = data.frame()
  
  year = NULL
  month = NULL
  
  m = 1 # month cursors
  sum = rep(0, k) # sum accumulator(row vector!)
  
  count = 0 # days accumulator
  ycount = 0 #  years accumulator
  for (i in (1:length(daily$months))) {
    # get all current values of snow, rain, temp, resp in vector
    curr = NULL
    for (j in (1:k)) {
      curr = c(curr, daily[[3+j]][i])
      # curr = c(curr, daily[i,(3+j)])
    }
    # curr = daily[i,][-(1:3)]: # this method is very slow!
    
    # sum: row vector
    # curr: row vector
    # avg: row vector
    # monthly: data frame/matrix
    
    # with the same month && not at December 31st
    if (m == daily$months[i] && i < length(daily$months)) {
      sum = sum + curr
      count = count + 1
    } # with the same month && at December 31st
    else if (m == daily$months[i] && i == length(daily$months)) {
      sum = sum + curr
      count = count + 1
      avg = sum / count
      monthly = rbind(monthly, avg)
      month = c(month, m)
      year = c(year, start+ycount)
    } # at the point of month change(1st day of next month)
    else {
      avg = sum / count
      monthly = rbind(monthly, avg)
      month = c(month, m)
      year = c(year, start+ycount)
      if (m == 12) {
        ycount = ycount + 1
      }
      m = daily$months[i]
      sum = curr
      count = 1
    }
  }
  # note k+3 = dim(daily)[2]
  names(monthly) = names(daily[4:(k+3)])
  return (data.frame(year, month, monthly))
}

```








<!-- ```{r, echo=TRUE} -->
<!-- start_time = Sys.time() -->
<!-- alldata = loadAllData("D:/NSERC/project/processedData/") -->
<!-- nd_time = Sys.time() -->
<!-- end_time - start_time -->


<!-- ``` -->










Operations by gridcodes
#==========================================================================================

```{r, echo=TRUE}
snow_fraction_by_gc = function(gridcode) {
  filename = paste0(gridcode, "_dat.RData")
  if (!(filename %in% FILENAMES)) {
    data = NULL
  } else {
    data = mget(load(paste0(PATH, filename)))
  }
  
  if (is.null(data)) {
    return (NA)
  }
  else {
    mts = mts(data)
    snow = mts$snow
    rain = mts$rain
    snow_sum = 0
    rain_sum = 0
    for (i in (1:length(snow))) {
      s = snow[i]
      r = rain[i]
      # both snow and rain have to be non NA!
      if (!is.na(s) && !is.na(r)) {
        snow_sum = snow_sum + s
        rain_sum = rain_sum + r
      }
    }
    snow_ratio = snow_sum / (snow_sum + rain_sum)
    return (snow_ratio)
  }
}

rain_fraction_by_gc = function(gridcode) {
  return (1 - snow_fraction_by_gc(gridcode))
}

```





```{r, echo=TRUE}
valid_gridcode = NULL
for (fn in FILENAMES) {
  n = nchar(fn)
  prefix = substr(fn, start=1, stop=n-10)
  valid_gridcode = c(valid_gridcode, as.numeric(prefix))
}

```

```{r, echo=TRUE}
get_fraction_by_gridcode = function(gridcodes) {
  snow_fraction = NULL
  rain_fraction = NULL
  gridcode = NULL
  for (gc in gridcodes) {
    snow_fraction = c(snow_fraction, snow_fraction_by_gc(gc))
    rain_fraction = c(rain_fraction, rain_fraction_by_gc(gc))
  }
  return (data.frame(gridcodes, snow_fraction, rain_fraction))
  
}


```

```{r, echo=TRUE}
start_time = Sys.time()
valid_snow_fraction = get_fraction_by_gridcode(valid_gridcode)
end_time = Sys.time()
end_time - start_time

```



```{r, echo=TRUE}

valid_watersheds = load_watersheds(gridcodes = valid_gridcode)
valid_snow_fraction = get_fraction(valid_watersheds)
```


```{r, echo=TRUE}
low_snow_gridcode = valid_snow_fraction$gridcodes[valid_snow_fraction$snow_fraction <= 0.03]


```



